---
title: "CNVworkflow Tables"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Supplmentary Table 1. Purity and Ploidy
```{r echo=FALSE}
data_dir = "~/data2/PureCN_manuscript/Figures/Final_Tables"
table_1 = read.csv(file.path(data_dir, "Table1_puri_ploi.csv"))[,-1]
```

* Total 675 samples (233 OV + 442 LUAD)
```{r echo=FALSE, eval=FALSE}
dim(table_1)
sum(table_1$Sample == "OV")
sum(table_1$Sample == "LUAD")
```

```{r}
head(table_1, 3)
```

## Supplmentary Table 2. LOH
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript/Figures/Final_Tables"
table_2 = read.csv(file.path(data_dir, "Table2_LOH.csv"))[,-1]
```

* Total 675 samples (233 OV + 442 LUAD)
```{r echo=FALSE, eval=FALSE}
dim(table_2)
sum(table_2$Sample == "OV")
sum(table_2$Sample == "LUAD")
```

```{r}
names(table_2)
head(table_2, 3)
```

## Supplmentary Table 3. AUC
* 233 OVC + 441 OVC samples         
  + 1 LUAD sample is absent in Table3 --> "TCGA-49-AARR-01"
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript/Figures/Final_Tables"
table_3 = read.csv(file.path(data_dir, "Table3_AUC.csv"))[,-1]
table_3
```

## Supplmentary Table 4. Benchmark
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript/Figures/Final_Tables"
table_4.1 = read.csv(file.path(data_dir, "Table4.1_benchmark_PCN.csv"))[,-1]
table_4.2 = read.csv(file.path(data_dir, "Table4.2_benchmark_SGZ.csv"))[,-1]
```

#### Table 4.1 Benchmark_PureCN
* Total 674 samples in Table4.1 (233 OV + 441 LUAD)     
    + 1 LUAD sample is absent in Table4.1 --> "TCGA-49-AARR-01"
```{r echo=FALSE, eval=FALSE}
dim(table_4.1)
sum(table_4.1$Sample == "OV")
sum(table_4.1$Sample == "LUAD")
```

```{r}
head(table_4.1, 3)
```

#### Table 4.2 Benchmark_SGZ
* Total 669 samples in Table4.2 (231 OV + 438 LUAD)      
  + 2 OVC samples are absent in Table4.2 --> "TCGA-61-1919-01" "TCGA-29-1691-01"   
  + 4 LUAD sampes are absent in Table4.2 --> "TCGA-05-4417-01" "TCGA-49-4514-01" "TCGA-55-8506-01" "TCGA-62-A46Y-01"
```{r echo=FALSE, eval=FALSE}
dim(table_4.2)
sum(table_4.2$Sample == "OV")
sum(table_4.2$Sample == "LUAD")
```

```{r}
head(table_4.2, 3)
```

## Supplmentary Table 5. Mutational Signatures
* Total 675 samples (233 OV + 442 LUAD)
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript/Figures/Final_Tables"
table_5 = read.csv(file.path(data_dir, "Table5_TMB.csv"))[,-1]
```

```{r}
head(table_5)
```

```{r}
tail(table_5)
```

## Supplmentary Table 6. Mutational Signatures
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript/Figures/Final_Tables"
table_6 = read.csv(file.path(data_dir, "Table6_mutSig.csv"))[,-1]

OVC_mutSig = table_6[which(table_6$Sample == "OV"),]
LUAD_mutSig = table_6[which(table_6$Sample == "LUAD"),]
```

* Total 466 OVC samples (233 Tumor-only + 233 Paired)
```{r echo=FALSE, eval=FALSE}
sum(OVC_mutSig$Mode == "Tumor-only")
sum(OVC_mutSig$Mode == "Paired")
```

```{r}
head(table_6, 3)
```

* Total 882 LUAD samples (442 Tumor-only + 440 Paired)     
  + 2 paired-mode LUAD samples are absent in Table4.2 --> "TCGA-49-AARR-01" "TCGA-86-A4P8-01" --> **Not enough somatic calls to deconstruct signatures.**
```{r echo=FALSE, eval=FALSE}
sum(LUAD_mutSig$Mode == "Tumor-only")
sum(LUAD_mutSig$Mode == "Paired")
```

```{r}
tail(table_6, 3)
```

## Supplmentary Table 7. Significant Genes from dNdSCV
From `~/data2/PureCN_manuscript/Revision/dNdSCV/script/dNdSCV_results.Rmd`

```{r}
table_7 = read.table(file.path(data_dir, "Table7_dNdSCV_SignificantGenes.csv"), sep = ",", header = TRUE)
dim(table_7)
head(table_7)
```

