---
title: "CNVworkflow Tables"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Supplmentary Table 1. Purity and Ploidy
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Tables"
table_1 = read.csv(file.path(data_dir, "Table1_puri_ploi.csv"))[,-1]
```

* Total 675 samples (233 OV + 442 LUAD)
```{r echo=FALSE, eval=FALSE}
dim(table_1)
sum(table_1$Sample == "OV")
sum(table_1$Sample == "LUAD")
```

```{r}
head(table_1, 3)
```

## Supplmentary Table 2. AUC
* 233 OVC + 441 OVC samples         
  + 1 LUAD sample is absent in Table3 --> "TCGA-49-AARR-01"
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Tables"
table_2 = read.csv(file.path(data_dir, "Table2_AUC.csv"))[,-1]
table_2
```

## Supplmentary Table 3. Benchmark
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Tables"
table_3.1 = read.csv(file.path(data_dir, "Table3.1_benchmark_PCN.csv"))[,-1]
table_3.2 = read.csv(file.path(data_dir, "Table3.2_benchmark_SGZ.csv"))[,-1]
```

#### Table 3.1 Benchmark_PureCN
* Total 674 samples in Table3.1 (233 OV + 441 LUAD)     
    + 1 LUAD sample is absent in Table3.1 --> "TCGA-49-AARR-01"
```{r echo=FALSE, eval=FALSE}
dim(table_3.1)
sum(table_3.1$Sample == "OV")
sum(table_3.1$Sample == "LUAD")
```

```{r}
head(table_3.1, 3)
```

#### Table 3.2 Benchmark_SGZ
* Total 669 samples in Table3.2 (231 OV + 438 LUAD)      
  + 2 OVC samples are absent in Table3.2 --> "TCGA-61-1919-01" "TCGA-29-1691-01"   
  + 4 LUAD sampes are absent in Table3.2 --> "TCGA-05-4417-01" "TCGA-49-4514-01" "TCGA-55-8506-01" "TCGA-62-A46Y-01"
```{r echo=FALSE, eval=FALSE}
dim(table_3.2)
sum(table_3.2$Sample == "OV")
sum(table_3.2$Sample == "LUAD")
```

```{r}
head(table_3.2, 3)
```

## Supplmentary Table 4. Mutational Signatures
```{r echo=FALSE}
data_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Tables"
table_4 = read.csv(file.path(data_dir, "Table4_mutSig.csv"))[,-1]

OVC_mutSig = table_4[which(table_4$Sample == "OV"),]
LUAD_mutSig = table_4[which(table_4$Sample == "LUAD"),]
```

* Total 466 OVC samples (233 Tumor-only + 233 Paired)
```{r echo=FALSE, eval=FALSE}
sum(OVC_mutSig$Mode == "Tumor-only")
sum(OVC_mutSig$Mode == "Paired")
```

```{r}
head(OVC_mutSig, 3)
```

* Total 882 LUAD samples (442 Tumor-only + 440 Paired)     
  + 2 paired-mode LUAD samples are absent in Table4.2 --> "TCGA-49-AARR-01" "TCGA-86-A4P8-01" --> **Not enough somatic calls to deconstruct signatures.**
```{r echo=FALSE, eval=FALSE}
sum(LUAD_mutSig$Mode == "Tumor-only")
sum(LUAD_mutSig$Mode == "Paired")
```

```{r}
head(LUAD_mutSig, 3)
```
