---
title: "EDA: mutation counts"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
})
```

```{r}
library(dplyr)
```

## Load the mutation counts
These mutation counts were extracted from 'somatic.ontarget' column of '_mutation_burden.csv' PureCN output files
```{r}
data_dir = "/home/sehyun/Documents/github/PureCN_Pipeline/analysis/mutation_burden"

catalog = "931070"
mut_count_931070 = read.table(file.path(data_dir, catalog, 
                               paste0(catalog, "_mutation_count")))

catalog = "S0293689"
mut_count_S0293689 = read.table(file.path(data_dir, catalog, 
                               paste0(catalog, "_mutation_count")))
```

## Merge duplicates
Merge mutational counts from both kits
```{r}
mut_count = rbind(mut_count_931070, mut_count_S0293689)
```

There are total 30 duplicated samples (25 analyzed by both kits + 5 analyzed from different analytes)   

##### 1. duplicates from both kits
```{r}
dup_kit = unlist(intersect(mut_count_931070$SampleId, 
                           mut_count_S0293689$SampleId))
dup_ind = which(mut_count$SampleId %in% dup_kit)
dup_subset = mut_count[dup_ind, ]
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 50 duplicates with 25 merged
```{r}
mut_count = mut_count[!mut_count$SampleId %in% dup_kit,]   # 307 rows --> 257 rows
mut_count = rbind(mut_count, dup_subset)   # 282 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_kit, dup_ind, dup_subset, catalog)
```

##### 2. duplicates from different analytes
```{r}
dup_analyte = read.csv("~/Documents/github/PureCN_manuscript/Results/dup_analyte.csv")$x
dup_ind = which(mut_count$SampleId %in% dup_analyte)
dup_subset = as.data.frame(lapply(mut_count[dup_ind, ], unlist))
```

Merge SampleId with analyte 'D' to that with analyte 'W'
```{r}
dup_subset$SampleId[c(1,3,5,7,9)] = dup_subset$SampleId[c(2,4,6,8,10)]
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 10 duplicates with 5 merged
```{r}
mut_count = mut_count[!mut_count$SampleId %in% dup_analyte,]   # 282 rows --> 272 rows
mut_count = rbind(mut_count, dup_subset)   # 277 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_analyte, dup_ind, dup_subset)
```

##### Save
```{r eval=FALSE}
mutSig_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results/mutSig"
write.table(mut_count, 
            file = file.path(mutSig_dir, "extdata/mut_count.tsv"))
```