---
title: "EAD: mutational signature from Dx.R output"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
})
```

```{r}
library(dplyr)
library(stringr)
```

# Tumor_only mode 
```{r}
run_type = "tumor_only"
```

## Capture kit 931070
##### Make a list with paths to all '_signature.csv' files
```{r}
catalog = "931070"
methods_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Methods"
source(file.path(methods_dir, "input_info.R"))

Dx_dir = file.path("/data/16tb/CNVworkflow", catalog, "purecn_output", paste0(catalog, "_Dx"))
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Switch full filenames with barcode
```{r}
# load full filename and barcode sampleMap
sampleMap = read.csv("/data/16tb/CNVworkflow/931070/sampleMap_931070.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(sig_summary)) {
  sig_summary[i, 1] = sampleMap[sampleMap$SampleId == sig_summary[i, 1], ]$analyte
}

sig_t_only_931070 = sig_summary
```

```{r echo=FALSE}
# clean the environment
rm(sig_summary, bed, catalog, kit, cds_dir, Dx_dir, fname, sig_file, sampleMap, i, methods_dir)
```

## Capture kit S0293689
##### Make a list with paths to all '_signature.csv' files
```{r}
catalog = "S0293689"
methods_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Methods"
source(file.path(methods_dir, "input_info.R"))

Dx_dir = file.path("/data/16tb/CNVworkflow", catalog, "purecn_output", paste0(catalog, "_Dx"))
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Switch full filenames with barcode
```{r}
# load full filename and barcode sampleMap
sampleMap = read.csv("/data/16tb/CNVworkflow/S0293689/sampleMap_S0293689.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(sig_summary)) {
  sig_summary[i, 1] = sampleMap[sampleMap$SampleId == sig_summary[i, 1], ]$analyte
}

sig_t_only_S0293689 = sig_summary
```

```{r echo=FALSE, warning=FALSE}
# clean the environment
rm(sig_summary, bed, catalog, kit, cds_dir, Dx_dir, fname, sig_file, sampleMap, i, methods_dir)
```

## Merge duplicates
Merge mutational signature outputs from both kits with tumor_only samples
```{r}
dat = rbind(sig_t_only_931070, sig_t_only_S0293689)   # 324 obs.
```

```{r echo=FALSE, eval=FALSE}
data_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results/mutSig/data"
tumor_only = saveRDS(dat, file.path(data_dir, "sig_all_tumor_only_2.rds"))
```

There are total 32 duplicated samples (27 analyzed by both kits + 5 analyzed from different analytes)   

##### 1. duplicates from both kits
```{r}
dup_kit = read.csv("~/Documents/github/PureCN_manuscript/Results/dup_kit.csv")$x
dup_ind = which(dat$SampleId %in% dup_kit)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 54 duplicates with 27 merged
```{r}
dat = dat[!dat$SampleId %in% dup_kit,]   # 324 rows --> 270 rows
dat = rbind(dat, dup_subset)   # 297 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_kit, dup_ind, dup_subset)
```

##### 2. duplicates from different analytes
```{r}
dup_analyte = read.csv("~/Documents/github/PureCN_manuscript/Results/dup_analyte.csv")$x
dup_ind = which(dat$SampleId %in% dup_analyte)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))
```

Merge SampleId with analyte 'D' to that with analyte 'W'
```{r}
dup_subset$SampleId[c(1,3,5,7,9)] = dup_subset$SampleId[c(2,4,6,8,10)]
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 10 duplicates with 5 merged
```{r}
dat = dat[!dat$SampleId %in% dup_analyte,]   # 297 rows --> 287 rows
dat = rbind(dat, dup_subset)   # 292 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_analyte, dup_ind, dup_subset)
```

##### Save
```{r eval=FALSE}
results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results"
saveRDS(dat, 
        file.path(results_dir, "mutSig/data/sig_all_tumor_only.rds"))
```

# Matching_normal mode 
```{r}
run_type = "matching_normal"
```

## Capture kit 931070
##### Make a list with paths to all '_signature.csv' files
```{r}
catalog = "931070"
methods_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Methods"
source(file.path(methods_dir, "input_info.R"))

Dx_dir = file.path("/data/16tb/CNVworkflow", catalog, "purecn_output", paste0(catalog, "_Dx"))
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Switch full filenames with barcode
```{r}
# load full filename and barcode sampleMap
sampleMap = read.csv("/data/16tb/CNVworkflow/931070/sampleMap_931070.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(sig_summary)) {
  sig_summary[i, 1] = sampleMap[sampleMap$SampleId == sig_summary[i, 1], ]$analyte
}

sig_matching_normal_931070 = sig_summary
```

```{r echo=FALSE}
# clean the environment
rm(sig_summary, bed, catalog, kit, cds_dir, Dx_dir, fname, sig_file, sampleMap, i, methods_dir)
```

## Capture kit S0293689
##### Make a list with paths to all '_signature.csv' files
```{r}
catalog = "S0293689"
methods_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Methods"
source(file.path(methods_dir, "input_info.R"))

Dx_dir = file.path("/data/16tb/CNVworkflow", catalog, "purecn_output", paste0(catalog, "_Dx"))
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Switch full filenames with barcode
```{r}
# load full filename and barcode sampleMap
sampleMap = read.csv("/data/16tb/CNVworkflow/S0293689/sampleMap_S0293689.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(sig_summary)) {
  sig_summary[i, 1] = sampleMap[sampleMap$SampleId == sig_summary[i, 1], ]$analyte
}

sig_matching_normal_S0293689 = sig_summary
```

```{r echo=FALSE, warning=FALSE}
# clean the environment
rm(sig_summary, bed, catalog, kit, cds_dir, Dx_dir, fname, sig_file, sampleMap, i, methods_dir)
```

## Merge duplicates
Merge mutational signature outputs from both kits with matching_normal samples
```{r}
dat = rbind(sig_matching_normal_931070, sig_matching_normal_S0293689)
```

There are total 30 duplicated samples (25 analyzed by both kits + 5 analyzed from different analytes)   

##### 1. duplicates from both kits
```{r}
dup_kit = unlist(intersect(sig_matching_normal_931070$SampleId, 
                           sig_matching_normal_S0293689$SampleId))
dup_ind = which(dat$SampleId %in% dup_kit)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 50 duplicates with 25 merged
```{r}
dat = dat[!dat$SampleId %in% dup_kit,]   # 307 rows --> 257 rows
dat = rbind(dat, dup_subset)   # 282 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_kit, dup_ind, dup_subset)
```

##### 2. duplicates from different analytes
```{r}
dup_analyte = read.csv("~/Documents/github/PureCN_manuscript/Results/dup_analyte.csv")$x
dup_ind = which(dat$SampleId %in% dup_analyte)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))
```

Merge SampleId with analyte 'D' to that with analyte 'W'
```{r}
dup_subset$SampleId[c(1,3,5,7,9)] = dup_subset$SampleId[c(2,4,6,8,10)]
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 10 duplicates with 5 merged
```{r}
dat = dat[!dat$SampleId %in% dup_analyte,]   # 282 rows --> 272 rows
dat = rbind(dat, dup_subset)   # 277 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup_analyte, dup_ind, dup_subset)
```

##### Save
```{r eval=FALSE}
results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results"
saveRDS(dat, 
        file.path(results_dir, "mutSig/data/sig_all_matching_normal.rds"))
```