---
title: "dNdSCV outputs from tumor-only and matching-normal samples"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    toc_depth: 4
vignette: >
  % \VignetteIndexEntry{Statistical analysis output}
  % \VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE)
```


```{r eval=FALSE}
# Check KRAS 
# Extract KRAS status

i = 1
runtype = c("tumor_only", "matching_normal")[i]
data_dir = file.path("/nobackup/16tb_a/CNVworkflow_LUAD/purecn_output/PureCN", runtype)

sample_dir = list.files(data_dir)
file_list = sapply(sample_dir, function(x) list.files(file.path(data_dir, x)))
mut = sapply(file_list, function(x) x[grep("*_variants.csv$", x)])
source("~/Documents/github/PureCN_manuscript/Figures/Final_Figures/non_dup.R")
sub_ind = which(stringr::str_extract(names(mut), "TCGA.{11}") %in% luad_442)
mut = mut[sub_ind]

p = read.csv(file.path(data_dir, names(mut[i]), mut[[i]]))
colNames = colnames(p)
test = as.data.frame(matrix(NA, ncol = length(colNames)))
colnames(test) = colNames

for (i in seq_along(mut)) {
    p = read.csv(file.path(data_dir, names(mut[i]), mut[[i]]))
    q = p[which(p$gene.symbol == "KRAS"),]
    test = rbind(test, q)
}

# write.csv(test, paste0("~/data2/PureCN_manuscript/Revision/dNdSCV/data/KRAS_",runtype,".csv"))
```

```{r eval=FALSE}
out_dir = "~/data2/PureCN_manuscript/Revision/dNdSCV"
kras_t = read.csv(file.path(out_dir, "data/KRAS_tumor_only.csv"))[-1,-1]
kras_m = read.csv(file.path(out_dir, "data/KRAS_matching_normal.csv"))[-1,-1]
```




# Rescaling and Filtering
**Significant genes in tumor-only**   
Because I didn't use COSMIC VCF fils (COSMIC.CNT) for PureCN run, many SNVs in dbSNP 
were removed in the prior.somatic filtering process. Here, as a 'quick and dirty' fix,
we tried rescaling (`~/data2/PureCN_manuscript/Revision/dNdSCV/script/rescale_priors.R`). 


```{r eval=TRUE}
# FDR filtering
runtype = "tumor_only"
dnds_out_dir = file.path("/nobackup/16tb_b/CNVworkflow_revision/dNdSCV/outputs/LUAD", runtype)
# t_all = readRDS(file.path(dnds_out_dir, paste0("dndsout_LUAD_", runtype, "_rescalePriors.rds")))
t_all = readRDS(file.path(dnds_out_dir, paste0("dndsout_LUAD_", runtype, "_rescalePriors_PoNFiltered.rds")))
runtype = "matching_normal"
dnds_out_dir = file.path("/nobackup/16tb_b/CNVworkflow_revision/dNdSCV/outputs/LUAD", runtype)
m_all = readRDS(file.path(dnds_out_dir,"ML_SOMATIC_FILTER", paste0("dndsout_LUAD_", runtype, ".rds")))

t = t_all$sel_cv[,c("gene_name", "qglobal_cv")]
m = m_all$sel_cv[,c("gene_name", "qglobal_cv")]
```

```{r eval=TRUE}
# signif_dir = "/nobackup/16tb_b/CNVworkflow_revision/dNdSCV/outputs/LUAD"
# t = read.csv(file.path(signif_dir, "tumor_only/signif_genes_LUAD_tumor_only_rescalePriors.csv"))[,-1]
# m = read.csv(file.path(signif_dir, "matching_normal/ML_SOMATIC_FILTER/signif_genes_LUAD_matching_normal.csv"))[,-1]

colnames(t)[2] = "qglobal_cv_tumorOnly"
colnames(m)[2] = "qglobal_cv_matchingN"

all = merge(t, m, by = "gene_name")
# write.csv(all, "~/data2/PureCN_manuscript/Revision/dNdSCV/data/qgloba_cv_all.csv")
```

# Summary Table
## All the significant genes
Passed the FDR < 0.01 from tumor-only or mathing-normal analyses.
```{r message=FALSE, eval=TRUE}
# summary = dplyr::full_join(t, m)
# colnames(summary) = c("Gene", "Tumor only", "Matching normal")
# summary

FDRcutoff = 0.01
sub = all[which(all$qglobal_cv_tumorOnly < FDRcutoff | all$qglobal_cv_matchingN < FDRcutoff),]
colnames(sub) = c("Gene", "Tumor only", "Matching normal")
```

```{r echo=FALSE, eval=TRUE}
sub = sub[,c("Gene", "Matching normal", "Tumor only")]
sub = sub[order(sub$`Matching normal`),]
sub
```

```{r eval=FALSE}
# write.csv(sub, file = "~/data2/PureCN_manuscript/Revision/dNdSCV/data/SignificantGenes.csv")
write.csv(sub, file = "~/data2/PureCN_manuscript/Figures/Final_Tables/Table5_dNdSCV_SignificantGenes_PoNFiltered.csv", row.names = FALSE)
```

## Mutually called significant genes
Passed the FDR < 0.01 from **both** tumor-only and mathing-normal analyses.

```{r eval=TRUE}
FDRcutoff = 0.01
t_sig = t[which(t$qglobal_cv_tumorOnly < FDRcutoff),]
m_sig = m[which(t$qglobal_cv_tumorOnly < FDRcutoff),]
dplyr::inner_join(t_sig, m_sig, by = "gene_name")
all_sig = merge(t_sig, m_sig, by = "gene_name")
```

# Venn Diagram
```{r fig.width=4.5, fig.height=4.5, eval=TRUE}
library(VennDiagram)

grid.newpage()
draw.pairwise.venn(area1 = nrow(m_sig),
                   area2 = nrow(t_sig),
                   cross.area = nrow(all_sig),
                   category = c("matching_normal", "tumor_only"),
                   lty = rep("blank", 2), 
                   fill = c("light blue", "pink"), 
                   alpha = rep(0.5, 2), 
                   cat.pos = c(0, 0), 
                   cat.dist = rep(0.025, 2))
```

# Cancer Genes
```{r fig.height=4.5, fig.width=4.5, eval=TRUE}
plot(x = -log(all_sig$qglobal_cv_matchingN), 
     y = -log(all_sig$qglobal_cv_tumorOnly),
     main = "Cancer Genes in LUAD (-log(q-value))",
     xlab = "Matching-normal",
     ylab = "Tumor-only",
     xlim = c(0,20), ylim = c(0,20))
abline(0, 1, col = "red")
```

