---
title: "EDA: benchmark_SGZ"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
})
```

```{r}
library(dplyr)
library(stringr)
```

Data cleaning process here is referenced from 'EDA_mutational_signature.Rmd'   

## Load outputs from BenchmarkTumorOnly.R
```{r}
out.dir_931070 = "/data/16tb/CNVworkflow/931070/purecn_output/931070_Benchmark/tumor_only_SGZ"
output_931070 = list.files(out.dir_931070, pattern = ".csv")
import_931070 = file.path(out.dir_931070, output_931070)

benchmark_931070 = sapply(seq_along(import_931070), function(i) {
  read.csv(import_931070[i], stringsAsFactors = FALSE)
}) 

benchmark_931070= as.data.frame(t(benchmark_931070))
names(benchmark_931070)[1] = "SampleId"

# Switch full filenames with barcode
sampleMap = read.csv("/data/16tb/CNVworkflow/931070/sampleMap_931070.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(benchmark_931070)) {
  benchmark_931070[i, 1] = sampleMap[sampleMap$SampleId == benchmark_931070[i, 1], ]$analyte
}
```

```{r}
out.dir_S0293689 = "/data/16tb/CNVworkflow/S0293689/purecn_output/S0293689_Benchmark/tumor_only_SGZ"
output_S0293689 = list.files(out.dir_S0293689, pattern = ".csv")
import_S0293689 = file.path(out.dir_S0293689, output_S0293689)

benchmark_S0293689 = sapply(seq_along(import_S0293689), function(i) {
  read.csv(import_S0293689[i], stringsAsFactors = FALSE)
}) 

benchmark_S0293689= as.data.frame(t(benchmark_S0293689))
names(benchmark_S0293689)[1] = "SampleId"

# Switch full filenames with barcode
sampleMap = read.csv("/data/16tb/CNVworkflow/S0293689/sampleMap_S0293689.csv")
sampleMap$SampleId = gsub(".bam", "", sampleMap$filename)
sampleMap$analyte = str_extract(sampleMap$barcode, "TCGA.{16}")

for (i in 1:nrow(benchmark_S0293689)) {
  benchmark_S0293689[i, 1] = sampleMap[sampleMap$SampleId == benchmark_S0293689[i, 1], ]$analyte
}
```

## Merge duplicates
```{r}
dat = rbind(benchmark_931070, benchmark_S0293689)   # 303 obs.
```

There are total 25 duplicated samples
##### 1. duplicates from both kits
```{r}
dup_kit = unlist(dat[duplicated(dat$SampleId),]$SampleId)   # 24 obs.
dup_ind = which(dat$SampleId %in% dup_kit)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))   # 48 obs.
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.) 
```

Switch 48 duplicates with 24 merged
```{r}
dat = dat[!dat$SampleId %in% dup_kit,]   # 303 obs. --> 255 obs.
dat = rbind(dat, dup_subset)   # 279 obs.
```

##### 2. duplicates from different analytes
```{r}
dup_analyte = read.csv("~/Documents/github/PureCN_manuscript/Results/dup_analyte.csv")$x
dup_ind = which(dat$SampleId %in% dup_analyte)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))   # 10 obs.
```

Merge SampleId with analyte 'D' to that with analyte 'W'
```{r}
dup_subset$SampleId[c(1,3,5,7,9)] = dup_subset$SampleId[c(2,4,6,8,10)]
```

Merge duplicates by taking a mean of them
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.)
```

Switch 10 duplicates with 5 merged
```{r}
dat = dat[!dat$SampleId %in% dup_analyte,]   # 279 rows --> 269 rows
dat = rbind(dat, dup_subset)   # 274 x 8
```


```{r eval=FALSE}
results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results"
saveRDS(dat, 
        file.path(results_dir, "benchmark/data/all_ovc_benchmark_SGZ.rds"))
```
