---
title: "EDA: Purity & Ploidy from PureCN outputs"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(ggplot2)
  library(cowplot)
})
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(cowplot)
```

## Import purity/ploidy from ABSOLUTE
557 ABSOLUTE outputs are loaded as 'abs_puri_ploi' with the SampleId format, TCGA-XX-XXXX-XX
```{r}
script_dir = "/home/sehyun/Documents/github/PureCN_manuscript"
methods_dir = file.path(script_dir, "Methods")
results_dir = file.path(script_dir, "Results/purity_ploidy")

source(file.path(methods_dir, "sample_sorting.R"))
source(file.path(results_dir, "R/puri_ploi_function.R"))
source(file.path(results_dir, "R/purity_ploidy_ABS.R"))
```

## Import purity/ploidy from PureCN
### Tumor_only PureCN result
```{r}
purecn_mode = "tumor_only"

kit="931070"
source(file.path(results_dir, "R/purity_ploidy_import.R"))
res_tumor_only_931070 = res_all
rm(res, res_all)

kit="S0293689"
source(file.path(results_dir, "R/purity_ploidy_import.R"))
res_tumor_only_S0293689 = res_all
rm(res, res_all)

tumor_only = rbind(res_tumor_only_931070, res_tumor_only_S0293689)
tumor_only = as.data.frame(lapply(tumor_only, unlist))
# saveRDS(puri_ploi_tumor_only, file = file.path(results_dir, "data/puri_ploi_tumor_only.rds"))
```

### Matching_normal PureCN result
```{r}
purecn_mode = "matching_normal"

kit="931070"
source(file.path(results_dir, "R/purity_ploidy_import.R"))
res_matching_normal_931070 = res_all
rm(res, res_all)

kit="S0293689"
source(file.path(results_dir, "R/purity_ploidy_import.R"))
res_matching_normal_S0293689 = res_all
rm(res, res_all)

matching_normal = rbind(res_matching_normal_931070, res_matching_normal_S0293689)
matching_normal = as.data.frame(lapply(matching_normal, unlist))
# saveRDS(puri_ploi_matching_normal, file = file.path(results_dir, "data/puri_ploi_matching_normal.rds"))
```

## Combine ABSOLUTE and PureCN outputs (Tumor_only)
##### Check duplicates
- 27 samples are duplicated because they were analyzed by both capture kits   
- 5 samples are duplicated because they were analyzed by one kit (931070) from two different analyte types   
```{r}
dup_submitter = tumor_only[which(duplicated(tumor_only$submitter_id)),]$submitter_id
dup_subset = tumor_only[which(tumor_only$submitter_id %in% dup_submitter),]
dup_subset = dup_subset[order(dup_subset$submitter_id),]   # 64 x 7
dup_subset
```

```{r echo=FALSE}
t_dup = dup_subset
write.csv(t_dup, file = file.path(script_dir, "Results/duplication/ovc_t_dup.csv"))
```

```{r}
kit_dup_931070 = dup_subset[which(dup_subset$capture_kit == "931070"),][,c(1:3)]
kit_dup_S0293689 = dup_subset[which(dup_subset$capture_kit == "S0293689"),][,c(1:3)]
kit_dup_all = left_join(kit_dup_931070, kit_dup_S0293689, by = "Sampleid", suffix = c("_931070", "_S0293689"))
head(kit_dup_all)
```

**Purity comparison of 27 technical duplicates**

```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = kit_dup_all
x = "Purity_931070"
y = "Purity_S0293689"
plot_for = "purity"

customLabs = labs(x = "Purity from kit 931070", 
                  y = "Purity from kit S0293689", 
                  title = "Purity estimates from different capture kit") 

source("~/Documents/github/PureCN_manuscript/Figures/manuscript_figure.R")
```

**Purity comparison of 27 technical duplicates**   
The major outlier is 'TCGA-23-1118-01A-01W'.
```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = kit_dup_all
x = "Ploidy_931070"
y = "Ploidy_S0293689"
plot_for = "ploidy"

customLabs = labs(x = "Ploidy from kit 931070", 
                  y = "Ploidy from kit S0293689", 
                  title = "Ploidy estimates from different capture kit") 

source("~/Documents/github/PureCN_manuscript/Figures/manuscript_figure.R")
```

```{r}
sample_dup = c("TCGA-25-2396-01", "TCGA-25-2397-01", "TCGA-25-2398-01", "TCGA-59-2348-01", "TCGA-09-2044-01")
sample_dup = dup_subset[which(dup_subset$submitter_id %in% sample_dup),][,c(1:3, 6)]
sample_dup
```

##### Merge the duplicated entries
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(submitter_id) %>%
  dplyr::summarise(Sampleid = last(Sampleid),
            Purity = mean(Purity),
            Ploidy = mean(Ploidy),
            Flagged = paste(Flagged, collapse = '|'),
            Comment = paste(Comment, collapse = '|'),
            capture_kit = paste(capture_kit, collapse = '|')) %>%
  as.data.frame(.)
```

##### Merged duplicates    
- grouped by 'submitter_id'   
- If Sampleid from a submitter_id was different, selected the analyte 'W' over 'D'   
- Purity and Ploidy are merged by 'mean'   
- Flagged/ Coomment/ capture_kit info was merged with separater '|'   
```{r}
dup_subset
```

Replace 64 duplicates with 32 merged entries
```{r}
tumor_only = tumor_only[-which(tumor_only$submitter_id %in% dup_submitter),]
tumor_only = rbind(tumor_only, dup_subset) 

head(tumor_only, 3)
```

Rename some columns to combine with ABSOLUTE outputs   
- Sampleid (TCGA-XX-XXXX-XXX-XXX) --> fullname   
- submitter_id (TCGA-XX-XXXX-XX) --> SampleId   
- Purity --> Purity_tumor_only   
- Ploidy --> Ploidy_tumor_only
```{r}
names(tumor_only)[which(names(tumor_only) == "Sampleid")] = "fullname"
names(tumor_only)[which(names(tumor_only) == "submitter_id")] = "SampleId"
names(tumor_only)[which(names(tumor_only) == "Purity")] = "Purity_tumor_only"
names(tumor_only)[which(names(tumor_only) == "Ploidy")] = "Ploidy_tumor_only"
```

##### Merge
Combine them through 'SampleId'   
- among 292 PureCN outputs, 15 doesn't exist in ABSOLUTE outputs   
- among 550 ABSOLUTE outputs, 273 doesn't exist in PureCN outputs   
--> there are 277 common samples analyzed by both methods
```{r warning=FALSE}
ABS_w_tumor_only = left_join(abs_puri_ploi, tumor_only, by = "SampleId")   # 550 x 9
ABS_w_tumor_only = ABS_w_tumor_only[!is.na(ABS_w_tumor_only$fullname), ]   # 277 x 9
```

##### Save 
ABSOLUTE and tumor_only (PureCN) output
```{r eval=FALSE}
saveRDS(ABS_w_tumor_only, file = file.path(results_dir,"data/ABS_w_tumor_only.rds"))
```

## Combine ABSOLUTE and PureCN outputs (Mathing_normal)
##### Check duplicates
- 25 samples are duplicated because they were analyzed by both capture kits   
- 5 samples are duplicated because they were analyzed by one kit (931070) from two different analyte types   
```{r}
dup_submitter = matching_normal[which(duplicated(matching_normal$submitter_id)),]$submitter_id
dup_subset = matching_normal[which(matching_normal$submitter_id %in% dup_submitter),]
dup_subset = dup_subset[order(dup_subset$submitter_id),]   # 60 x 7
dup_subset
```

```{r echo=FALSE}
paired_dup = dup_subset
write.csv(paired_dup, file = file.path(script_dir, "Results/duplication/ovc_paired_dup.csv"))
```

##### Merge the duplicated entries
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(submitter_id) %>%
  dplyr::summarise(Sampleid = last(Sampleid),
            Purity = mean(Purity),
            Ploidy = mean(Ploidy),
            Flagged = paste(Flagged, collapse = '|'),
            Comment = paste(Comment, collapse = '|'),
            capture_kit = paste(capture_kit, collapse = '|')) %>%
  as.data.frame(.)
```

##### Merged duplicates    
- grouped by 'submitter_id'   
- If Sampleid from a submitter_id was different, selected the analyte 'W' over 'D'   
- Purity and Ploidy are merged by 'mean'   
- Flagged/ Coomment/ capture_kit info was merged with separater '|'   
```{r}
dup_subset
```

Replace 60 duplicates with 30 merged entries
```{r}
matching_normal = matching_normal[-which(matching_normal$submitter_id %in% dup_submitter),]
matching_normal = rbind(matching_normal, dup_subset) 

head(matching_normal, 3)
```

Rename some columns to combine with ABSOLUTE outputs   
- Sampleid (TCGA-XX-XXXX-XXX-XXX) --> fullname   
- submitter_id (TCGA-XX-XXXX-XX) --> SampleId   
- Purity --> Purity_matching_normal   
- Ploidy --> Ploidy_matching_normal
```{r}
names(matching_normal)[which(names(matching_normal) == "Sampleid")] = "fullname"
names(matching_normal)[which(names(matching_normal) == "submitter_id")] = "SampleId"
names(matching_normal)[which(names(matching_normal) == "Purity")] = "Purity_matching_normal"
names(matching_normal)[which(names(matching_normal) == "Ploidy")] = "Ploidy_matching_normal"
```

##### Merge
Combine them through 'SampleId' and remove the rows that we didn't process using PureCN + 12 rows that were processed by PureCN but not by ABSOLUTE
```{r warning=FALSE}
ABS_w_matching_normal = left_join(abs_puri_ploi, matching_normal, by = "SampleId")
ABS_w_matching_normal = ABS_w_matching_normal[!is.na(ABS_w_matching_normal$fullname), ]
```

##### Save 
ABSOLUTE and matching_normal (PureCN) output
```{r eval=FALSE}
saveRDS(ABS_w_matching_normal, 
        file.path(results_dir,"data/ABS_w_matching_normal.rds"))
```

## Combine PureCN outputs: with or without matching-normal
Using the 'fullname' (= barcode, TCGA-XX-XXXX-XXX-XXX)
```{r warning=FALSE}
purecn_output = left_join(tumor_only, matching_normal, 
                         by = "fullname", 
                         suffix = c("_tumor_only", "_matching_normal"))

purecn_output = purecn_output[!is.na(purecn_output$Purity_matching_normal),]
```

##### Save
merged
```{r eval=FALSE}
saveRDS(purecn_output, file = file.path(results_dir,"data/purecn_only.rds"))
```
