---
title: "EAD: LUAD mutational signature from Dx.R output"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
})
```

```{r}
library(dplyr)
library(stringr)
```

# Tumor_only mode 
```{r}
run_type = "tumor_only"
```

## Capture kit 931070
##### Make a list with paths to all '_signature.csv' files
```{r}
Dx_dir = "/data/ovarian/CNVworkflow_LUAD/purecn_output/Dx"
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Add submitter_id column to group data
```{r}
dat = sig_summary
```

```{r echo=FALSE}
# clean the environment
rm(sig_summary, cds_dir, Dx_dir, fname, sig_file, run_type)
```

```{r eval=FALSE, echo=FALSE}
luad_442 = read.csv("~/Documents/github/PureCN_manuscript/luad/Results/duplication/luad_442.csv")[,2]
dat = dat[which(stringr::str_extract(dat$SampleId, "TCGA.{11}") %in% stringr::str_extract(luad_442, "TCGA.{11}")),]

results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/luad/Results"
saveRDS(dat, 
        file.path(results_dir, "mutSig/data/luad_sig_all_tumor_only_2.rds"))
```

## Merge duplicates
26 patients' samples are multiplicated generating total 84 different data points.

```{r}
dup = read.csv("~/Documents/github/PureCN_manuscript/luad/Results/dup_subset.csv")
dup_ind = which(stringr::str_extract(dat$SampleId, "TCGA.{16}") %in% dup$Sampleid)
dup_subset = as.data.frame(lapply(dat[dup_ind, ], unlist))
```

Merge duplicates by taking a mean of them
```{r}
dup_subset$SampleId = stringr::str_extract(dup_subset$SampleId, "TCGA.{11}")
dup_subset = dup_subset %>%
  dplyr::group_by(SampleId) %>%
  dplyr::summarise_all(funs(mean)) %>%
  as.data.frame(.) 
```

Switch 84 multi-plicates with 26 merged
```{r}
dat = dat[!stringr::str_extract(dat$SampleId, "TCGA.{11}") %in% dup$submitter_id,]   # 567 rows --> 483 rows
dat = rbind(dat, dup_subset)   # 509 x 31
```

```{r echo=FALSE}
# clean the environment
rm(dup, dup_ind, dup_subset)
```

##### Save
```{r eval=FALSE}
results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/luad/Results"
saveRDS(dat, 
        file.path(results_dir, "mutSig/data/luad_sig_all_tumor_only.rds"))
```

# Matching-normal mode 
```{r}
run_type = "matching_normal"
```

## Capture kit 931070
##### Make a list with paths to all '_signature.csv' files
```{r}
Dx_dir = "/data/ovarian/CNVworkflow_LUAD/purecn_output/Dx"
cds_dir = file.path(Dx_dir, paste0(run_type, "_cds"))
fname = list.files(cds_dir)

sig_file = file.path(cds_dir, fname, paste0(fname, "_signatures.csv"))
sig_file = sig_file[file.exists(sig_file)]
```

##### Import all mutational signature
Make a data frame with all the mutational signatures for each sample
```{r}
sig_summary = sapply(seq_along(sig_file), function(i) {
  read.csv(sig_file[i], stringsAsFactors = FALSE)
}) 

sig_summary = as.data.frame(t(sig_summary))
names(sig_summary)[1] = "SampleId"
```

Add submitter_id column to group data
```{r}
dat = sig_summary
```

```{r echo=FALSE}
# clean the environment
rm(sig_summary, cds_dir, Dx_dir, fname, sig_file, run_type)
```

```{r eval=FALSE, echo=FALSE}
results_dir = "/home/sehyun/Documents/github/PureCN_manuscript/luad/Results"
saveRDS(dat, 
        file.path(results_dir, "mutSig/data/luad_sig_all_matching_normal.rds"))
```