---
title: "EDA: TCGA-LUAD Purity & Ploidy from PureCN outputs"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(ggplot2)
})
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
```

## Import purity/ploidy from ABSOLUTE
507 ABSOLUTE outputs are loaded as 'abs_puri_ploi' (out of 509 overlapping submitters, 2 didn't have a output)
```{r message=FALSE, warning=FALSE}
script_dir = "/home/sehyun/Documents/github/PureCN_manuscript"
results_dir = file.path(script_dir, "luad/Results/purity_ploidy")
source(file.path(results_dir, "R/luad_purity_ploidy_ABS.R"))
```

## Import purity/ploidy from PureCN
### Tumor_only PureCN result
```{r}
source(file.path(script_dir, "Results/purity_ploidy/R/puri_ploi_function.R"))
```

```{r}
purecn_mode = "tumor_only"

kit="931070"
source(file.path(results_dir, "R/luad_purity_ploidy_import.R"))

tumor_only = res_all   # 567 obs.
tumor_only = as.data.frame(lapply(tumor_only, unlist))
```

```{r}
saveRDS(LUAD_puri_ploi_all, file = file.path(results_dir,"data/luad_puri_ploi_all_2.rds"))
```


### Matching-normal PureCN result
```{r}
purecn_mode = "matching_normal"

kit="931070"
source(file.path(results_dir, "R/luad_purity_ploidy_import.R"))

matching_normal = res_all   # 444 obs.
matching_normal = as.data.frame(lapply(matching_normal, unlist))
```

## Combine ABSOLUTE and PureCN outputs (Tumor_only)
##### Check duplicates
```{r}
dup_submitter = unique(tumor_only[which(duplicated(tumor_only$submitter_id)),]$submitter_id)   # 26
dup_subset = tumor_only[which(tumor_only$submitter_id %in% dup_submitter),]
dup_subset = as.data.frame(lapply(dup_subset, unlist))
dup_subset = dup_subset[order(dup_subset$submitter_id),]   # 84 x 7
dup_subset
```

```{r eval=FALSE, echo=FALSE}
write.csv(dup_subset, file = "~/Documents/github/PureCN_manuscript/luad/Results/duplication/dup_subset.csv")
```

##### Merge the duplicated entries
```{r}
dup_subset = dup_subset %>%
  dplyr::group_by(submitter_id) %>%
  dplyr::summarise(Sampleid = last(Sampleid),
            Purity = mean(Purity),
            Ploidy = mean(Ploidy),
            Flagged = paste(Flagged, collapse = '|'),
            Comment = paste(Comment, collapse = '|'),
            capture_kit = paste(capture_kit, collapse = '|')) %>%
  as.data.frame(.)
```

##### Merged duplicates    
- grouped by 'submitter_id'   
- If Sampleid from a submitter_id was different, selected the analyte 'W' over 'D'   
- Purity and Ploidy are merged by 'mean'   
- Flagged/ Coomment/ capture_kit info was merged with separater '|'   
```{r}
dup_subset   # 26 obs.
```

Replace 64 duplicates with 32 merged entries
```{r}
tumor_only = tumor_only[-which(tumor_only$submitter_id %in% dup_submitter),]
tumor_only = rbind(tumor_only, dup_subset)   # 509 obs. = 567 - 84 + 26

head(tumor_only, 3)
```

Rename some columns to combine with ABSOLUTE outputs   
- Sampleid (TCGA-XX-XXXX-XXX-XXX) --> fullname   
- submitter_id (TCGA-XX-XXXX-XX) --> SampleId   
- Purity --> Purity_tumor_only   
- Ploidy --> Ploidy_tumor_only
```{r}
names(tumor_only)[which(names(tumor_only) == "Sampleid")] = "fullname"
names(tumor_only)[which(names(tumor_only) == "submitter_id")] = "SampleId"
names(tumor_only)[which(names(tumor_only) == "Purity")] = "Purity_tumor_only"
names(tumor_only)[which(names(tumor_only) == "Ploidy")] = "Ploidy_tumor_only"
```

##### Merge
Combine them through 'SampleId'   
- among 509 PureCN outputs, 1 doesn't exist in ABSOLUTE outputs   
--> there are 507 common samples analyzed by both methods
```{r warning=FALSE}
ABS_w_tumor_only = dplyr::left_join(abs_puri_ploi, tumor_only, by = "SampleId")   # 508 x 9
ABS_w_tumor_only = ABS_w_tumor_only[!is.na(ABS_w_tumor_only$fullname), ]   # 507 x 9
```

##### Save 
ABSOLUTE and tumor_only (PureCN) output
```{r eval=FALSE}
saveRDS(ABS_w_tumor_only, file = file.path(results_dir,"data/luad_ABS_w_tumor_only.rds"))
```

```{r echo=FALSE, eval=FALSE}
luad_ABS_w_tumor_only = readRDS(file.path(results_dir, "data/luad_ABS_w_tumor_only.rds"))
res = luad_ABS_w_tumor_only[-which(luad_ABS_w_tumor_only$SampleId %in% dup_submitter),]
res = res[,c(4,2,3,5,6)]
names(res)[4:5] = c("Purity_PureCN", "Ploidy_PureCN")
write.table(res, file = file.path(results_dir, "extdata/LUAD_puri_ploi.csv"), row.names = FALSE)

# res = read.csv(file.path(results_dir, "extdata/LUAD_puri_ploi.csv"), sep = " ")
```

## Combine ABSOLUTE and PureCN outputs (Matching-normal)
```{r echo=FALSE}
names(matching_normal)[which(names(matching_normal) == "Sampleid")] = "fullname"
names(matching_normal)[which(names(matching_normal) == "submitter_id")] = "SampleId"
names(matching_normal)[which(names(matching_normal) == "Purity")] = "Purity_matching_normal"
names(matching_normal)[which(names(matching_normal) == "Ploidy")] = "Ploidy_matching_normal"
```

##### Merge
Combine them through 'SampleId'   
- among 444 PureCN (matching_normal) outputs, 2 doesn't exist in ABSOLUTE outputs   
--> there are 442 common samples analyzed by both methods
```{r warning=FALSE}
ABS_w_matching_normal = dplyr::left_join(abs_puri_ploi, matching_normal, by = "SampleId")   # 507 x 9
ABS_w_matching_normal = ABS_w_matching_normal[!is.na(ABS_w_matching_normal$fullname), ]   # 442 x 9
```

##### Save 
ABSOLUTE and matching_normal (PureCN) output
```{r eval=FALSE}
saveRDS(ABS_w_matching_normal, file = file.path(results_dir,"data/luad_ABS_w_matching_normal.rds"))
```

```{r echo=FALSE, eval=FALSE}
## ABS + tumor_only + matching_normal 
abs_tumor_only = ABS_w_tumor_only[,c("SampleId", "fullname", "Purity_ABS", "Ploidy_ABS", "Purity_tumor_only", "Ploidy_tumor_only")]  # 507 obs.
matching_normal_only = ABS_w_matching_normal[,c("SampleId", "Purity_matching_normal", "Ploidy_matching_normal")]  # 442 obs.
LUAD_puri_ploi_all = dplyr::left_join(abs_tumor_only, matching_normal_only, by = "SampleId")
LUAD_puri_ploi_all = LUAD_puri_ploi_all[!is.na(LUAD_puri_ploi_all$Purity_matching_normal), ]   # 442 obs.
saveRDS(LUAD_puri_ploi_all, file = file.path(results_dir,"data/luad_puri_ploi_all.rds"))
```

