---
title: "EDA: duplicated samples"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TCGA-OV
```{r}
manifest = readRDS("~/Documents/github/PureCN_manuscript/Data/ovc_manifest/ovc_manifest_112618.rds")

# Subset manifest with only one bedfile
bedfiles = manifest$bedfiles
summary = data.frame()
for (i in 1:length(bedfiles)){
  summary = rbind(summary, c(i, length(bedfiles[[i]])))
  colnames(summary) = c("i", "num_bedfiles")
}

ovc_mani = manifest[which(summary$num_bedfiles == 1),]
ovc_mani = ovc_mani[which(ovc_mani$target_capture_kit_catalog_number %in% c("931070", "S0293689")),]
rm(bedfiles, manifest, summary, i)
```

```{r}
ovc_mani = ovc_mani[, c("barcode", "submitter_id", "sample", "analyte", "target_capture_kit_catalog_number")]   # 660 obs.
```

```{r}
ovc_t = ovc_mani[which(ovc_mani$sample == 1),]    # class is 'interger'   # 327 obs.
ovc_n = ovc_mani[which(ovc_mani$sample != 1),]    # 333 obs.
```

```{r}
# the number of samples for ABS vs. PureCN (tumor_only) after merging
length(unique(ovc_t$submitter_id))   # 295 obs.

# the number of samples for tumor_only vs. matched_normal after merging
length(intersect(unique(ovc_t$submitter_id), unique(ovc_n$submitter_id)))   # 280 obs.
```

246 non-duplicated submitters
```{r message=FALSE}
library(dplyr)
dup_t = ovc_t %>% group_by(submitter_id) %>% filter(n()>1)   # 64 obs.
ovc_t = ovc_t[which(!ovc_t$barcode %in% dup_t$barcode),]   # 263 obs. <-- 327 obs.
dup_n = ovc_n %>% group_by(submitter_id) %>% filter(n()>1)   # 62 obs.
ovc_n = ovc_n[which(!ovc_n$barcode %in% dup_n$barcode),]   # 271 obs. <-- 333 obs.

matching = which(ovc_t$submitter_id %in% ovc_n$submitter_id)
bed_subset = ovc_t[matching,]   # 312 obs. (246 obs. after the above filtering)
```

557 ABSOLUTE outputs are loaded as 'abs_puri_ploi' with the SampleId format, TCGA-XX-XXXX-XX
```{r}
script_dir = "/home/sehyun/Documents/github/PureCN_manuscript"
methods_dir = file.path(script_dir, "Methods")
results_dir = file.path(script_dir, "Results/purity_ploidy")

source(file.path(methods_dir, "sample_sorting.R"))
source(file.path(results_dir, "R/puri_ploi_function.R"))
source(file.path(results_dir, "R/purity_ploidy_ABS.R"))
```

236 non-duplicated submitters have ABSOLUTE output
```{r}
x = intersect(abs_puri_ploi$SampleId, stringr::str_extract(bed_subset$barcode, "TCGA.{11}"))
length(x)
write.csv(x, "~/Documents/github/PureCN_manuscript/Results/duplication/ovc_236.csv")
```

```{r echo=FALSE, eval=FALSE}
# applying all the filtering in the above chunk --> 246 obs.
bed_subset_fullname = stringr::str_extract(bed_subset$barcode, "TCGA.{16}")
write.csv(bed_subset_fullname, "~/Documents/github/PureCN_manuscript/Results/duplication/ovc_246.csv")
```

```{r message=FALSE, warning=FALSE, eval=FALSE, echo=FALSE}
data_dir = "~/Documents/github/PureCN_manuscript/Results/purity_ploidy"
purecn_only = readRDS(file.path(data_dir, "data/purecn_only.rds"))
purecn_only = purecn_only[,c("fullname", "Purity_tumor_only", "Ploidy_tumor_only", "SampleId_tumor_only", "capture_kit_tumor_only", "Purity_matching_normal", "Ploidy_matching_normal", "capture_kit_matching_normal")]
purecn_only$submitter_id = stringr::str_extract(purecn_only$fullname, "TCGA.{8}")

x = ovc_t[which(ovc_t$barcode %in% dup_t$barcode),] %>% as.data.frame(.) 
x$fullname = stringr::str_extract(x$barcode, "TCGA.{16}")

x = cbind(x, purecn_only, by = "submitter_id")
x[order(x$submitter_id),]
```

```{r}
ovc_paired = data.frame()
for (i in 1:nrow(bed_subset)) {
  barcode <- bed_subset[i, "barcode"]
  submitter_id <- bed_subset[i, "submitter_id"]
  sample <- bed_subset[i, "sample"]
  analyte <- bed_subset[i, "analyte"]
  kit <- bed_subset[i, "target_capture_kit_catalog_number"]
  
  matching_i <- which(ovc_n$submitter_id == submitter_id)
  matching_barcode <- ovc_n[matching_i, "barcode"]
  matching_submitter_id <- ovc_n[matching_i, "submitter_id"]
  matching_sample <- ovc_n[matching_i, "sample"]
  matching_analyte <- ovc_n[matching_i, "analyte"]
  matching_kit <- ovc_n[matching_i, "target_capture_kit_catalog_number"]

  ovc_paired_i = data.frame(barcode, submitter_id, sample, analyte, kit, 
                            matching_barcode, matching_submitter_id, matching_sample, matching_analyte, matching_kit)
  ovc_paired = rbind(ovc_paired, ovc_paired_i)
}
```

```{r echo=FALSE}
rm(analyte, barcode, i, kit, matching, matching_analyte, matching_barcode, matching_i, matching_kit, matching_sample, matching_submitter_id, sample, submitter_id, ovc_paired_i)
```

### the number of unique submitter for TCGA-OV tumor and normal
```{r}
length(unique(ovc_t$submitter_id))   # 295 out of 327
length(unique(ovc_n$submitter_id))   # 302 out of 333
```

### tumor/normal pair from different analyte types
```{r}
sum(ovc_paired$analyte != ovc_paired$matching_analyte)
# ovc_paired[which(ovc_paired$analyte != ovc_paired$matching_analyte),]
```

### tumor/normal pair from different capture kit
```{r}
sum(ovc_paired$kit != ovc_paired$matching_kit)
# ovc_paired[which(ovc_paired$kit != ovc_paired$matching_kit),]
```

### analytes used
```{r}
table(ovc_n$analyte)
table(ovc_t$analyte)
```

## TCGA-LUAD
```{r message=FALSE}
luad_mani = readRDS("~/Documents/github/PureCN_manuscript/luad/luad_manifest_annot.rds")
luad_download = readr::read_tsv("/data/ovarian/luad_manifest_download_1.tsv")
luad_mani = luad_mani[which(luad_mani$id %in% luad_download$id),]
rm(luad_download)
```

```{r}
luad_mani = luad_mani[, c("barcode", "submitter_id", "sample", "analyte", "target_capture_kit_catalog_number")]   # 1162 obs.
```

```{r}
luad_t = luad_mani[which(luad_mani$sample %in% c("01", "02")),]   # 568 obs. # class is still 'character'
luad_n = luad_mani[which(!luad_mani$sample %in% c("01", "02")),]   # 594 obs.
```

```{r}
# the number of samples for ABS vs. PureCN (tumor_only) after merging
length(unique(luad_t$submitter_id))   # 507 obs.

# the number of samples for tumor_only vs. matched_normal after merging
length(intersect(unique(luad_t$submitter_id), unique(luad_n$submitter_id)))   # 506 obs.
```

```{r message=FALSE}
library(dplyr)
dup_t = luad_t %>% group_by(submitter_id) %>% filter(n()>1)   # 89 obs.
luad_t = luad_t[which(!luad_t$barcode %in% dup_t$barcode),]   # 479 obs. <-- 568 obs.
dup_n = luad_n %>% group_by(submitter_id) %>% filter(n()>1)   # 144 obs.
luad_n = luad_n[which(!luad_n$barcode %in% dup_n$barcode),]   # 450 obs. <-- 594 obs.

matching = which(luad_t$submitter_id %in% luad_n$submitter_id)
bed_subset = luad_t[matching,]   # 444 obs.
```

```{r}
luad_paired = data.frame()
for (i in 1:nrow(bed_subset)) {
  barcode <- bed_subset[i, "barcode"]
  submitter_id <- bed_subset[i, "submitter_id"]
  sample <- bed_subset[i, "sample"]
  analyte <- bed_subset[i, "analyte"]
  kit <- bed_subset[i, "target_capture_kit_catalog_number"]
  
  matching_i <- which(luad_n$submitter_id == submitter_id)
  matching_barcode <- luad_n[matching_i, "barcode"]
  matching_submitter_id <- luad_n[matching_i, "submitter_id"]
  matching_sample <- luad_n[matching_i, "sample"]
  matching_analyte <- luad_n[matching_i, "analyte"]
  matching_kit <- luad_n[matching_i, "target_capture_kit_catalog_number"]

  luad_paired_i = data.frame(barcode, submitter_id, sample, analyte, kit, 
                            matching_barcode, matching_submitter_id, matching_sample, matching_analyte, matching_kit)
  luad_paired = rbind(luad_paired, luad_paired_i)
}
```

```{r echo=FALSE}
rm(analyte, barcode, i, kit, matching, matching_analyte, matching_barcode, matching_i, matching_kit, matching_sample, matching_submitter_id, sample, submitter_id, luad_paired_i)
```

### the number of unique submitter for TCGA-OV tumor and normal
```{r}
length(unique(luad_t$submitter_id))   # 479 out of 479
length(unique(luad_n$submitter_id))   # 450 out of 450
```

### tumor/normal pair from different analyte types
```{r}
sum(luad_paired$analyte != luad_paired$matching_analyte)
```

### analytes used
```{r}
table(luad_n$analyte)
table(luad_t$analyte)
```

