---
title: "CNVworkflow: Purity and Ploidy"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
})
```

```{r}
library(ggplot2)
library(cowplot)
```

```{r}
script_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup"
methods_dir = file.path(script_dir, "Methods")
results_dir = file.path(script_dir, "Results/purity_ploidy")
```

```{r echo=FALSE, eval=FALSE}
script_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup"
methods_dir = file.path(script_dir, "Methods")
results_dir = file.path(script_dir, "Results/purity_ploidy")

source(file.path(methods_dir, "sample_sorting.R"))
source(file.path(results_dir, "R/puri_ploi_function.R"))
source(file.path(results_dir, "R/purity_ploidy_ABS.R"))

puri_ploi_931070 = read.csv(file.path(results_dir, "extdata/puri_ploi_table_931070.csv"))
puri_ploi_S0293689 = read.csv(file.path(results_dir, "extdata/puri_ploi_table_S0293689.csv"))

puri_ploi_931070$kit = "931070"
puri_ploi_S0293689$kit = "S0293689"
purecn_puri_ploi = rbind(puri_ploi_931070[-1], puri_ploi_S0293689[-1])   # 323 obs.
```

## Tumor-only mode 
Results from 233 samples are plotted (out of 277 tumor-only processed samples)
```{r}
source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Figures/non_dup.R")
purecn_puri_ploi = readRDS(file.path(results_dir,"data/ABS_w_tumor_only.rds"))   # 277 obs.
purecn_puri_ploi = purecn_puri_ploi[which(purecn_puri_ploi$SampleId %in% ovc_236),]   # 233 obs.
```

```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
hist(abs(purecn_puri_ploi$Purity_ABS - purecn_puri_ploi$Purity_tumor_only), 
     xlab = "abs(ABS$purity - PureCN$purity)", 
     main = "Abs. diff. b/w purity (TCGA-OV tumor only)",
     ylim = c(0, 150))
```

```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
hist(abs(purecn_puri_ploi$Ploidy_ABS - purecn_puri_ploi$Ploidy_tumor_only), 
     xlab = "abs(ABS$ploidy - PureCN$ploidy)", 
     main = "Abs. diff. b/w ploidy (TCGA-OV tumor only)",
     ylim = c(0, 250))
```

### Purity Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = purecn_puri_ploi
x = "Purity_ABS"
y = "Purity_tumor_only"
plot_for = "purity"

customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "A. TCGA-OV Purity") 

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
tumor_puri = final_plot
rm(final_plot)

# out_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Figures"
# saveRDS(tumor_puri, file = file.path(out_dir, "Fig1A.ovc_puri.rds"))
```

### Ploidy Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = purecn_puri_ploi
x = "Ploidy_ABS"
y = "Ploidy_tumor_only"
plot_for = "ploidy"

customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "B. TCGA-OV Ploidy") 

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
tumor_ploi = final_plot
rm(final_plot)

# saveRDS(tumor_ploi, file = file.path(out_dir, "Fig1B.ovc_ploi.rds"))
```

## Matching-normal mode
Results from 233 samples are plotted
```{r}
puri_ploi_all = readRDS(file.path(results_dir,"data/ABS_w_matching_normal.rds"))   # 265 obs.
puri_ploi_all = puri_ploi_all[which(puri_ploi_all$SampleId %in% ovc_236),]   # 233 obs.

head(puri_ploi_all, 3)
```

### Purity Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = puri_ploi_all
x = "Purity_ABS"
y = "Purity_matching_normal"
plot_for = "purity"

customLabs = labs(x = "SNP6 array (ABSOLUTE)",
                  y = "Paired WES (PureCN)",
                  title = "TCGA-OV Purity")

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
paired_puri = final_plot
rm(final_plot)
```

### Ploidy Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df = puri_ploi_all
x = "Ploidy_ABS"
y = "Ploidy_matching_normal"
plot_for = "ploidy"

customLabs = labs(x = "SNP6 array (ABSOLUTE)",
       y = "Paired WES (PureCN)",
       title = "TCGA-OV Ploidy")

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
paired_ploi = final_plot
rm(final_plot)
```

Use `plot.margin = unit(c(0.5,0.9,0.3,0.7), "cm"))`...
in `~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure_style.R`
```{r fig.width = 8, fig.height = 9}
library(patchwork)
(tumor_puri | tumor_ploi) /
  (paired_puri | paired_ploi)
```

## Compare PureCN outputs: with or wihtout matching-normal 
Results from 307 samples are plotted
```{r eval=FALSE}
purecn_puri_ploi = readRDS(file.path(results_dir,"data/purecn_only.rds"))
```

### Purity Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5, eval=FALSE}
df = purecn_puri_ploi
x = "Purity_tumor_only"
y = "Purity_matching_normal"
plot_for = "purity"

customLabs = labs(x = "Tumor only", 
                  y = "Tumor paired", 
                  title = "A. TCGA-OV Purity") 

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
```

### Ploidy Comparisons
```{r echo=FALSE, fig.width = 5, fig.height = 4.5, eval=FALSE}
df = purecn_puri_ploi
x = "Ploidy_tumor_only"
y = "Ploidy_matching_normal"
plot_for = "ploidy"

customLabs = labs(x = "Tumor only", 
                  y = "Tumor paired", 
                  title = "B. TCGA-OV Ploidy") 

source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure.R")
```

```{r}
sessionInfo()
```

