---
title: "BenchmarkTumorOnly Plots - Both"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(cowplot)
  library(patchwork)
  library(stringr)
})
```

```{r}
library(dplyr)
library(cowplot)
library(patchwork)
library(stringr)
```

```{r echo=FALSE}
source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/manuscript_figure_style.R")
source("~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Figures/non_dup.R")
```

## Load OV benchmark output
```{r}
# PureCN algorithm
results_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Results"
benchmark = readRDS(file.path(results_dir, "benchmark/data/all_ovc_benchmark.rds"))

for (i in 2:10) {
  benchmark[, i] = as.numeric(benchmark[, i])
}

# SGZ algorithm
benchmark_SGZ = readRDS(file.path(results_dir, "benchmark/data/all_ovc_benchmark_SGZ.rds"))

for (i in 2:8) {
  benchmark_SGZ[, i] = as.numeric(benchmark_SGZ[, i])
}

# Combine both
benchmark$algorithm = "PureCN"
benchmark_SGZ$algorithm = "SGZ"
merge_column = c("SampleId", "Purity", "Ploidy", "Median.Coverage", "Callrate", "algorithm")

benchmark_OV = rbind(benchmark[, merge_column], benchmark_SGZ[, merge_column])
benchmark_OV$Sample = "OV"

# SGZ failed to get the results from 2 samples
benchmark_OV = benchmark_OV[which(str_extract(benchmark_OV$SampleId, "TCGA.{11}") %in% ovc_236),]
```

##### Relationship of Callrate vs. Purity and Median Coverage 
*only with PureCN algorithm, not SGZ*
```{r}
summary(lm(Callrate ~ Purity + Median.Coverage, data = benchmark))
```

## Load LUAD benchmark output
```{r}
# PureCN algorithm
results_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/luad/Results"
benchmark = readRDS(file.path(results_dir, "benchmark/data/all_luad_benchmark.rds"))
for (i in 2:10) {
  benchmark[, i] = as.numeric(benchmark[, i])
}

# SGZ algorithm
benchmark_SGZ = readRDS(file.path(results_dir, "benchmark/data/all_luad_benchmark_SGZ.rds"))

for (i in 2:8) {
  benchmark_SGZ[, i] = as.numeric(benchmark_SGZ[, i])
}

# Combine both
benchmark$algorithm = "PureCN"
benchmark_SGZ$algorithm = "SGZ"
merge_column = c("SampleId", "Purity", "Ploidy", "Median.Coverage", "Callrate", "algorithm")

benchmark_LUAD = rbind(benchmark[, merge_column], benchmark_SGZ[, merge_column])
benchmark_LUAD$Sample = "LUAD"

# PureCN and SGZ failed to get the results from 1 and 4 samples, respectively.
benchmark_LUAD = benchmark_LUAD[which(str_extract(benchmark_LUAD$SampleId, "TCGA.{11}") %in% luad_442),]
```

##### Relationship of Callrate vs. Purity and Median Coverage 
*only with PureCN algorithm, not SGZ*
```{r}
summary(lm(Callrate ~ Purity + Median.Coverage, data = benchmark))
```

## Plot Purity vs. Callrate
```{r}
benchmark_all = rbind(benchmark_LUAD, benchmark_OV)
```

##### callrate_plot
```{r fig.width = 4.5, fig.height = 4.5}
benchmark = benchmark_all[which(benchmark_all$Sample == "OV"),]
ov_callrate_plot = ggplot(benchmark, aes(Purity, Callrate, col = algorithm)) +
  geom_point(alpha = 1/3) + 
  scale_color_manual(values=c("red", "navy")) +
  labs(x = "Purity",
       y = "Callrate", 
       title = "C. TCGA-OV") + 
  customPlot +
  theme(legend.position = c(0.16, 0.13),
        legend.title = element_blank(),
        legend.text = element_text(size = 12, face = "plain"),
        legend.background = element_rect(fill = NA)) +
  scale_x_continuous(limits=c(0, 1), breaks=seq(0,1,0.2)) +
  scale_y_continuous(limits=c(0, 1), breaks=seq(0,1,0.2))
```

##### SGZ_callrate_plot
```{r fig.width = 4.5, fig.height = 4.5}
benchmark = benchmark_all[which(benchmark_all$Sample == "LUAD"),]
luad_callrate_plot = ggplot(benchmark, aes(Purity, Callrate, col = algorithm)) +
  geom_point(alpha = 1/3) + 
  scale_color_manual(values=c("red", "navy")) +
  labs(x = "Purity",
       y = "Callrate", 
       title = "D. TCGA-LUAD") + 
  customPlot +
  theme(legend.position = c(0.16, 0.13),
        legend.title = element_blank(),
        legend.text = element_text(size = 12, face = "plain"),
        legend.background = element_rect(fill = NA)) +
  scale_x_continuous(limits=c(0, 1), breaks=seq(0,1,0.2)) +
  scale_y_continuous(limits=c(0, 1), breaks=seq(0,1,0.2))
```

```{r fig.width = 9, fig.height = 4.5}
ov_callrate_plot + luad_callrate_plot

# final_fig_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Figures"
# saveRDS(ov_callrate_plot, file = file.path(final_fig_dir, "Fig2C.ov_callrate_plot.rds"))
# saveRDS(luad_callrate_plot, file = file.path(final_fig_dir, "Fig2D.luad_callrate_plot.rds"))
```

## PureCN algorithm outputs
### Plot SNV calling accuracy using PureCN algorithm
```{r fig.width = 5, fig.height = 5.5}
rm(benchmark)
benchmark = benchmark_all[which(benchmark_all$algorithm == "PureCN"),]
cov_plot_purecn = ggplot(benchmark, aes(Median.Coverage, Callrate, col = Sample)) + 
  geom_point(alpha = 1/3) + 
  scale_color_manual(values=c("navy", "red")) +
  labs(title = "A. PureCN",
       x = "Median Coverage") + 
  customPlot +
  theme(legend.position = c(0.24, 0.78),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, face = "plain"),
        plot.margin = unit(c(0.5,0.5,0.2,0.2), "cm"),
        legend.background = element_rect(fill = NA)) +
  # scale_x_continuous(limits=c(0, 1), breaks=seq(0,1,0.2)) +
  # scale_y_continuous(limits=c(0, 1), breaks=seq(0,1,0.2))
  facet_wrap(~round(Purity, digits = 1))
```

## PureCN algorithm outputs
### Plot SNV calling accuracy using SGZ algorithm
```{r fig.width = 5, fig.height = 5.5}
rm(benchmark)
benchmark = benchmark_all[which(benchmark_all$algorithm == "SGZ"),]
cov_plot_sgz = ggplot(benchmark, aes(Median.Coverage, Callrate, col = Sample)) + 
  geom_point(alpha = 1/3) + 
  scale_color_manual(values=c("navy", "red")) +
  labs(title = "B. SGZ",
       x = "Median Coverage") + 
  customPlot +
  theme(legend.position = c(0.24, 0.78),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, face = "plain"),
        plot.margin = unit(c(0.5,0.5,0.2,0.2), "cm"),
        legend.background = element_rect(fill = NA)) +
  # scale_x_continuous(limits=c(0, 1), breaks=seq(0,1,0.2)) +
  # scale_y_continuous(limits=c(0, 1), breaks=seq(0,1,0.2))
  facet_wrap(~round(Purity, digits = 1))
```

```{r fig.width = 10, fig.height = 5.5}
cov_plot_purecn + cov_plot_sgz
 
final_fig_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/Final_Figures"
saveRDS(cov_plot_purecn, file = file.path(final_fig_dir, "SFig4A.cov_plot_purecn.rds"))
saveRDS(cov_plot_sgz, file = file.path(final_fig_dir, "SFig4B.cov_plot_sgz.rds"))
```

## Callrate summary between OV and LUAD for each algorithm
```{r echo=FALSE}
s1 = summary(benchmark_OV[which(benchmark_OV$algorithm == "PureCN"),]$Callrate)
s2 = summary(benchmark_LUAD[which(benchmark_LUAD$algorithm == "PureCN"),]$Callrate)
s3 = summary(benchmark_OV[which(benchmark_OV$algorithm == "SGZ"),]$Callrate)
s4 = summary(benchmark_LUAD[which(benchmark_LUAD$algorithm == "SGZ"),]$Callrate)

summary = rbind(s1, s2, s3, s4) %>% as.data.frame(.)
info = data.frame(Cancer = c("OV", "LUAD", "OV", "LUAD"), Algorithm = c("PureCN", "PureCN", "SGZ", "SGZ"))

bind_cols(info, summary)
```

```{r message=FALSE, echo=FALSE}
callrate_hist = ggplot(benchmark_all, aes(x = algorithm, y = Callrate, col = Sample)) +
  geom_boxplot() + 
  scale_color_manual(values=c("navy", "red")) +
  labs(title = "G. Callrate summary") + 
  theme_classic() +
  theme(legend.position = c(0.15, 0.12),
        legend.title = element_blank(),
        legend.text = element_text(size = 10, face = "plain"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 12, face = "bold"), 
        axis.text.y = element_text(size = 10, face = "bold"), 
        axis.title.x = element_blank(),
        axis.title.y = element_text("Callrate", size = 12, face = "bold"),
        plot.margin = unit(c(0.5,0.5,0.2,0.2), "cm"),
        legend.background = element_rect(fill = NA),
        panel.border = element_rect(colour = "grey", fill = NA, size = 0.5)
)
```

```{r eval=FALSE, fig.width = 4, fig.height = 4.5}
callrate_hist

# final_fig_dir = "~/wallabe4_backup/github/PureCN_manuscript_backup/Figures/final"
# saveRDS(callrate_hist, file = file.path(final_fig_dir, "Fig2G.callrate_hist.rds"))
```


```{r}
sessionInfo()
```