---
title: "EAD: TCGA-LUAD"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Load manifest file
```{r}
luad_dir = "~/Documents/github/PureCN_manuscript/luad"
manifest = readRDS(file.path(luad_dir, "luad_manifest_annot.rds"))

head(manifest, 3)
```

##### Downloaded files
All the TCGA-LUAD bam files are in `/data/ovarian/luad` directory    
(1162 bam files which have corresponding results from ABSOLUTE)
```{r warning=FALSE, message=FALSE}
luad_manifest_download = readr::read_tsv("/data/ovarian/luad_manifest_download_1.tsv")
manifest = manifest[manifest$id %in% luad_manifest_download$id,]
```

##### Normal vs. Tumor samples
```{r}
manifest_n = manifest[manifest$sample_definition %in% 
                        c("Blood Derived Normal", "Solid Tissue Normal"),]
manifest_t = manifest[manifest$sample_definition %in% 
                        c("Primary Solid Tumor", "Recurrent Solid Tumor"),]

dim(manifest_t)
dim(manifest_n)
```

##### Capture kit used   
- catalog = "931070"   
- kit = "Custom V2 Exome Bait, 48 RXN X 16 tubes"   
- bed = "whole_exome_agilent_1.1_refseq_plus_3_boosters.targetIntervals.bed"
```{r}
table(manifest$target_capture_kit_catalog_number)
```

##### Unique submitter_id
```{r}
length(unique(manifest_t$submitter_id))
length(unique(manifest_n$submitter_id))
```

There are 506 tumor samples with matching normal that can compared to ABSOLUTE
```{r}
sum(unique(manifest_t$submitter_id) %in% unique(manifest_n$submitter_id))
```

```{r luad_sample_sorting}
# All TCGA-LUAD samples were processed by cat# 931070
catalog = "931070"
source("~/Documents/github/PureCN_manuscript/Methods/input_info.R")

# Load manifest file
luad_dir = "~/Documents/github/PureCN_manuscript/luad"
manifest = readRDS(file.path(luad_dir, "luad_manifest_annot.rds"))

# Downloaded bam files which have corresponding ABSOLUTE results
luad_manifest_download = readr::read_tsv("/data/ovarian/luad_manifest_download_1.tsv")
manifest = manifest[manifest$id %in% luad_manifest_download$id,]
manifest$file_dir = "/data/ovarian/luad"
rm(luad_manifest_download)

# Separate manifest into normal(n) and tumor(t) samples
manifest_n = manifest[manifest$sample_definition %in% 
                        c("Blood Derived Normal", "Solid Tissue Normal"),]
manifest_t = manifest[manifest$sample_definition %in% 
                        c("Primary Solid Tumor", "Recurrent Solid Tumor"),]

kit_subset_t = manifest_t[which(manifest_t$target_capture_kit_name == kit),]
bed_subset_t = kit_subset_t[which(kit_subset_t$bedfiles == bed),]   # 568 obs.

kit_subset_n = manifest_n[which(manifest_n$target_capture_kit_name == kit),]
bed_subset_n = kit_subset_n[which(kit_subset_n$bedfiles == bed),]   # 594 obs.
```

```{r}
library(stringr)
library(dplyr)

dup = bed_subset_t %>% group_by(submitter_id) %>% filter(n()>1)   # 89 obs.
dup_t = bed_subset_t[bed_subset_t$barcode %in% dup$barcode,]
dup_t[order(dup_t$barcode),][,c("barcode", "submitter_id", "sample", "analyte")]
length(unique(str_extract(dup_t$barcode, "TCGA.{11}")))   # 568 - 89 + 30 = 509
```

```{r}
dup = bed_subset_n %>% group_by(submitter_id) %>% filter(n()>1)   # 144 obs.
dup_n = bed_subset_n[bed_subset_n$barcode %in% dup$barcode,]
dup_n[order(dup_n$barcode),][,c("barcode", "submitter_id", "sample", "analyte")]
length(unique(str_extract(dup_n$barcode, "TCGA.{11}")))   # 594 - 144 + 102 = 552
```
