---
title: "Annotate manifest file created by GenomicDataCommons (LUAD)"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Build an annotated manifest file using GenomicDataCommons and TCGAutils packages
### Download manifest file using GenomicDataCommons
```{r eval=FALSE}
library(TCGAutils)
library(jsonlite)
library(curl)
library(downloader)

# using GenomicDataCommons package to generate manifest file
library(GenomicDataCommons)
library(magrittr)

# available_values('files','cases.project.project_id')
# available_values('files','experimental_strategy')
# available_values('files','data_format')

manifest <- GenomicDataCommons::files() %>%
  GenomicDataCommons::filter(~ cases.project.project_id == "TCGA-LUAD" &
           experimental_strategy == "WXS" &
           data_format == "BAM") %>%
  GenomicDataCommons::manifest()

manifest_raw <- manifest
```

```{r eval=FALSE}
manifest = cbind(manifest, 
                 UUIDtoBarcode(manifest$id, id_type = "file_id", 
                               end_point = "center"))

names(manifest)[7] = "barcode"
manifest = cbind(manifest, TCGAutils::TCGAbiospec(manifest$barcode))

manifest_dir = "~/Documents/github/PureCN_manuscript/luad"
write.table(manifest, file = file.path(manifest_dir, "luad_manifest.tsv"), sep="\t")
```

### Load the saved manifest file
```{r eval=FALSE}
manifest = read.delim(file.path(manifest_dir, "luad_manifest.tsv"), as.is = TRUE)
```

### curl information about target_capture_kit 
1) target_region, 2) name, 3) vender, 4) catalog_number
```{r eval=FALSE}
res = lapply(manifest$id, function(uuid) {
  con = curl::curl(paste0("https://api.gdc.cancer.gov/files/", uuid, "?pretty=true&fields=analysis.metadata.read_groups.target_capture_kit_target_region,analysis.metadata.read_groups.target_capture_kit_name,analysis.metadata.read_groups.target_capture_kit_vendor,analysis.metadata.read_groups.target_capture_kit_catalog_number"))
  x = jsonlite::fromJSON(con)
  return(x)
})

y = lapply(res, function(x) unique(x$data$analysis$metadata$read_groups))
y = do.call(rbind, y)
manifest = cbind(manifest, y)
```

### Cleaning target_region (BED files) information
Seperate the name of bedfiles from 'target_capture_kit_target_region'
```{r eval=FALSE}
bedfiles = tail(unlist(strsplit(as.character(manifest$target_capture_kit_target_region)[1], 
                                split="/", fixed=TRUE)), n = 1)
unique(bedfiles)
```

```{r eval=FALSE}
manifest$bedfiles = bedfiles
```

```{r echo=FALSE, eval=FALSE}
saveRDS(manifest, file = file.path(manifest_dir, "luad_manifest_annot.rds"))
rm(manifest)
```

## Download BAM files using gdc-client
```{r}
manifest_luad = readRDS(file.path(manifest_dir, "luad_manifest_annot.rds"))

# import ABSOLUTE outputs
file.dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results/ABSOLUTE/ABSOLUTE_sampleMap"
ABSOLUTE = read.delim(file.path(file.dir, "ABSOLUTE_Purity_Ploidy.txt"))

# subset LUAD data analyzed with ABSOLUTE
overlap_luad = str_extract(manifest_luad$barcode, "TCGA.{8}") %in% str_extract(ABSOLUTE$array, "TCGA.{8}")
luad_manifest_download = manifest_luad[overlap_luad,]
manifest_raw = manifest_raw[manifest_raw$md5 %in% luad_manifest_download$md5, ]

write.table(manifest_raw, 
            file = "/data/ovarian/luad_manifest_download.tsv", 
            quote=FALSE, sep="\t", row.names=FALSE)
```

```{r eval=FALSE}
# restart the paused download
setwd("/data/ovarian/luad")
x = list.files(getwd())
manifest_sub = manifest_raw[!manifest_raw$id %in% x,]

write.table(manifest_sub, 
            file = "/data/ovarian/luad_manifest_download_2.tsv", 
            quote=FALSE, sep="\t", row.names=FALSE)
```

```{r echo=FALSE, eval=FALSE}
luad_manifest_download = read_tsv("/data/ovarian/luad_manifest_download.tsv")
```

need to use the most up-to-date version of GenomicDataCommons to generate working manifest files
```{bash eval=FALSE}
/home/sehyun/gdc-client download -m /data/ovarian/luad_manifest_download_2.tsv \
-t /data/ovarian/gdc-user-token.2018-10-15T20_38_37.020Z.txt
```

