---
title: "Build a sampleMap for ABSOLUTE"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=FALSE, message=FALSE}
library(stringr)
```

## Load the manifest file for microarray experiment downloaded from GDC
```{r}
file.dir = "/home/sehyun/Documents/github/PureCN_manuscript/Results/ABSOLUTE/ABSOLUTE_sampleMap"
manifest_cnv = read.delim(file.path(file.dir, "manifest_cnv.csv"), as.is = TRUE)
```

## Check the sample types --> use only tumor samples
##### GDC Sample Type Code
- 01: Primary Solid Tumor   
- 02: Recurrent Solid Tumor   
- 10: Blood Derived Normal   
- 11: Solid Tissue Normal   

```{r}
table(manifest_cnv$sample)
manifest_cnv_tumor = manifest_cnv[which(manifest_cnv$sample %in% c(1, 2)),]
```

##### Number of samples with different types of tumor
```{r eval=FALSE}
manifest_cnv_tumor_1 = manifest_cnv[which(manifest_cnv$sample == 1),]   # 1146 obs.
cnv_tumor_name_1 = unique(manifest_cnv_tumor_1$barcode)   # 573 obs.

manifest_cnv_tumor_2 = manifest_cnv[which(manifest_cnv$sample == 2),]   # 54 obs.
cnv_tumor_name_2 = unique(manifest_cnv_tumor_2$barcode)   # 27 obs.
```

##### Out of 1200 entries in OVC_CNV manifest, these are the number of unique samples at the different levels of resolution, labled in its name.
```{r}
unique_barcode = unique(manifest_cnv_tumor$barcode)   # 600
unique_analyte = unique(str_extract(unique_barcode, "TCGA.{16}"))   # 587
unique_sample = unique(str_extract(unique_barcode, "TCGA.{11}"))    # 586
unique_submitter = unique(str_extract(unique_barcode, "TCGA.{8}"))   # 570
```

## Subset OV data from the whole ABSOLUTE data
##### 'ABSOLUTE_OV' contains ABSOLUTE outputs subset based on the information (= unique_sample) from the OVC_CNV manifest downloaded from GDC
```{r}
ABSOLUTE = read.delim(file.path(file.dir, "ABSOLUTE_Purity_Ploidy.txt"))

OV_index = which(ABSOLUTE$array %in% unique_sample)   # 557
ABSOLUTE_OV = ABSOLUTE[OV_index,]
```

```{r}
head(ABSOLUTE_OV)
```

## Build the sampleMap connecting ABSOLUTE outputs and CNV manifest through 'array' (TCGA-XX-XXX-XX)
##### I used 'sample' level to combine ABSOLUTE outputs to WES data, because once I add 'analyte' information, too many ABSOLUTE outputs can't be matched to WES data. 
```{r warning=FALSE}
sampleMap = data.frame(fullname = unique_barcode)
sampleMap$array = str_extract(sampleMap$fullname, "TCGA.{11}")   # 'sample' level resolution
sampleMap = dplyr::left_join(ABSOLUTE_OV, sampleMap, by = "array")
```

##### Check the duplicattion from fullname
```{r}
dup_array = sampleMap[which(duplicated(sampleMap$array)),]$array
sampleMap[which(sampleMap$array %in% dup_array),]
```

##### Merge the duplicated entries
5 samples are duplicated at the array level because they are different at the OVC_CNV fullname level (= one array/sample with different plate/center)   
```{r warning=FALSE}
dup_subset = sampleMap[which(sampleMap$array %in% dup_array),]
dup_subset = dup_subset %>%
  dplyr::group_by(array) %>%
  dplyr::summarise(sample = first(sample),
            call.status = first(call.status),
            purity = first(purity),
            ploidy = first(ploidy),
            Genome.doublings = first(Genome.doublings),
            Coverage.for.80..power = first(Coverage.for.80..power),
            Cancer.DNA.fraction = first(Cancer.DNA.fraction),
            Subclonal.genome.fraction = first(Subclonal.genome.fraction),
            solution = first(solution),
            fullname = paste(fullname, collapse = '|')) %>%
  as.data.frame(.)
```

Merged duplicates (two fullnames were combined with '|')
```{r}
dup_subset
```

Replace 10 duplicates with 5 merged entries
```{r}
sampleMap = sampleMap[-which(sampleMap$array %in% dup_array),]
sampleMap = rbind(sampleMap, dup_subset)
```

Change colname from 'array' to 'SampleId'
```{r}
colnames(sampleMap)[1] = "SampleId"
```

#### Save 
```{r eval=FALSE}
saveRDS(sampleMap, file = file.path(file.dir, "ABSOLUTE_sampleMap.rds"))
```

## To compare ABSOLUTE output against PureCN output
```{r}
sampleMap = readRDS(file.path(file.dir, "ABSOLUTE_sampleMap.rds"))
head(sampleMap, 3)
```

```{r}
sampleMap = sampleMap[, c("SampleId", "purity", "ploidy", "Genome.doublings", "Subclonal.genome.fraction", "fullname")]
head(sampleMap, 3)
abs_puri_ploi = sampleMap
```

