---
title: "Revision: Purity and Ploidy Outliers"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    toc_depth: 4
vignette: >
  % \VignetteIndexEntry{Statistical analysis output}
  % \VignetteEngine{knitr::rmarkdown}
---

```{r setup, echo=FALSE, results="hide", warning=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(patchwork)
})
```

# OVC tumor-only mode
## Load data
```{r}
script_dir = "~/data2/PureCN_manuscript"
results_dir = file.path(script_dir, "Results/purity_ploidy")
purecn_puri_ploi = readRDS(file.path(results_dir, "data/ABS_w_tumor_only.rds"))

source("~/Documents/github/PureCN_manuscript/Figures/Final_Figures/non_dup.R")
purecn_puri_ploi = purecn_puri_ploi[which(purecn_puri_ploi$SampleId %in% ovc_236),]
```

## Ploidy outliers
```{r outlier_percentage, echo=FALSE, eval=FALSE}
df = purecn_puri_ploi
x = df$Ploidy_ABS
y = df$Ploidy_tumor_only

absdiff <- abs(x - y)
ind <- order(absdiff)
fract <- 0.8
length.out <- round(fract * length(absdiff))
fract.grid <- ind[seq_len(length.out)]
```

```{r outlier_cutoff, echo=FALSE}
df = purecn_puri_ploi
x = df$Ploidy_ABS
y = df$Ploidy_tumor_only

absdiff <- abs(x - y)
fract.grid <- which(absdiff < 0.5)
```

Ploidy-concordant samples with the absolute ploidy difference less than 0.5, are 
marked in red and the others are colored in cyan. Pearson correlation on upper-left
side is calculated only with the concordant samples.

```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df$absdiff = "discordant"
df$absdiff[fract.grid] = "concordant"
df_ploi = df

x = "Ploidy_ABS"
y = "Ploidy_tumor_only"
plot_for = "ploidy"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "B. TCGA-OV Ploidy") 

source("~/data2/PureCN_manuscript/Revision/revision_figure.R")
tumor_ploid = final_plot
rm(final_plot)

tumor_ploid
```


```{r echo=FALSE, fig.width = 7, fig.height = 4.5, eval=FALSE}
# ## Purity of ploidy outliers
# **[Left]** Purity of PureCN and ABSOLUTE results from ploidy-discordant samples 
# (the cyan dots from the above plot) are plotted. 
# **[Right]** Purity of PureCN and ABSOLUTE results from the ploidy-concordant samples.  

df = df[which(df$absdiff == "discordant"),]
x = "Purity_ABS"
y = "Purity_tumor_only"
plot_for = "purity"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "TCGA-OV Purity \n of discordant Ploidy")

source("~/data2/PureCN_manuscript/Figures/manuscript_figure.R")
discordant_puri = final_plot
rm(final_plot)

df = df_ploi
df = df[which(df$absdiff == "concordant"),]
x = "Purity_ABS"
y = "Purity_tumor_only"
plot_for = "purity"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "TCGA-OV Purity \n of cordant Ploidy")

source("~/data2/PureCN_manuscript/Figures/manuscript_figure.R")
concordant_puri = final_plot
rm(final_plot)

discordant_puri | concordant_puri
```

```{r import_OV_QC, echo=FALSE}
catalogs = c("931070", "S0293689")

for (catalog in catalogs) {
    out_dir = file.path("/nobackup/16tb_b/CNVworkflow/", catalog, "purecn_output")
    data_dir = file.path(out_dir, paste0(catalog, "_tumor_cov"))
    fname_all = list.files(data_dir)
    fname_ind = grep("_coverage_loess_qc.txt", fname_all)
    fname = fname_all[fname_ind]
    
    colNames = colnames(read.table(file.path(data_dir, fname[1]), header = TRUE))
    QC = as.data.frame(matrix(NA, ncol = length(colNames)+1))
    colnames(QC) = c("SampleId", colNames)
    
    for (x in fname) {
        p = read.table(file.path(data_dir, x), header = TRUE)
        p$SampleId = x
        QC = rbind(QC, p)
    }
    assign(paste0("QC_", catalog), QC[-1,])
}
```

```{r echo=FALSE}
library(magrittr)
p = QC_S0293689$SampleId %>% gsub("_coverage_loess_qc.txt", "", .)

# match filenames with sample names
mani = read.table("~/data2/PureCN_manuscript/Data/manifest/ovc_manifest.tsv")
x = as.data.frame(matrix(NA, nrow = length(p), ncol=3))
colnames(x) = c("file_name", "barcode", "sample_name")
x$file_name = p
for (i in 1:nrow(x)) {
    y = x$file_name[i]
    if (!identical(y, "character(0)")) {
        z = mani$barcode[which(gsub(".bam", "",mani$filename) == y)]
        x$barcode[i] = as.character(z)
        x$sample_name[i] = stringr::str_extract(x$barcode[i], "TCGA.{11}")
    } else {
        next
    }
}

QC_931070$SampleId = stringr::str_extract(QC_931070$SampleId, "TCGA.{11}")
QC_S0293689$SampleId = x$sample_name
```

```{r echo=FALSE}
QC_all = rbind(QC_931070, QC_S0293689)
OV_all = dplyr::inner_join(df, QC_all, by = "SampleId")
```

```{r echo=FALSE}
x = OV_all$Ploidy_ABS
y = OV_all$Ploidy_tumor_only

absdiff <- abs(x - y)
fract.grid <- which(absdiff < 0.5)

OV_all$absdiff = "discordant"
OV_all$absdiff[fract.grid] = "concordant"
```








# LUAD tumor-only mode
## Load data
```{r}
script_dir = "~/data2/PureCN_manuscript"
results_dir = file.path(script_dir, "luad/Results/purity_ploidy")
purecn_puri_ploi = readRDS(file.path(results_dir, "data/luad_ABS_w_tumor_only.rds"))

# 442 obs. which have matched normal samples
# paired = readRDS(file.path(results_dir, "data/luad_ABS_w_matching_normal.rds"))
# write.csv(paired$fullname, "~/data2/PureCN_manuscript/luad/Results/duplication/luad_442.csv")
luad_442 = read.csv("~/data2/PureCN_manuscript/luad/Results/duplication/luad_442.csv")[,2]  

purecn_puri_ploi = purecn_puri_ploi[which(purecn_puri_ploi$fullname %in% luad_442),]
```

## Ploidy outliers
```{r echo=FALSE, eval=FALSE}
df = purecn_puri_ploi
x = df$Ploidy_ABS
y = df$Ploidy_tumor_only

absdiff <- abs(x - y)
ind <- order(absdiff)
fract <- 0.8
length.out <- round(fract * length(absdiff))
fract.grid <- ind[seq_len(length.out)]
# plot(x[fract.grid], y[fract.grid])
```

```{r echo=FALSE}
df = purecn_puri_ploi
x = df$Ploidy_ABS
y = df$Ploidy_tumor_only

absdiff <- abs(x - y)
fract.grid <- which(absdiff < 0.5)
```


Ploidy-concordant samples with the absolute ploidy difference less than 0.5, are 
marked in red and the others are colored in cyan. Pearson correlation on upper-left
side is calculated only with the concordant samples.

```{r echo=FALSE, fig.width = 5, fig.height = 4.5}
df$absdiff = "discordant"
df$absdiff[fract.grid] = "concordant"
df_ploi = df

x = "Ploidy_ABS"
y = "Ploidy_tumor_only"
plot_for = "ploidy"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "D. TCGA-LUAD Ploidy") 

source("~/data2/PureCN_manuscript/Revision/revision_figure.R")
tumor_ploid = final_plot
rm(final_plot)

tumor_ploid
```


```{r import_LUAD_QC, echo=FALSE}
data_dir = "/nobackup/16tb_a/CNVworkflow_LUAD/purecn_output/tumor_cov"
fname_all = list.files(data_dir)
fname_ind = grep("_coverage_loess_qc.txt", fname_all)
fname = fname_all[fname_ind]

colNames = colnames(read.table(file.path(data_dir, fname[1]), header = TRUE))
QC = as.data.frame(matrix(NA, ncol = length(colNames)+1))
colnames(QC) = c("SampleId", colNames)

for (x in fname) {
  p = read.table(file.path(data_dir, x), header = TRUE)
  p$SampleId = stringr::str_extract(x, "TCGA.{11}")
  QC = rbind(QC, p)
}

QC = QC[-1,]
```

```{r echo=FALSE}
LUAD_all = dplyr::inner_join(df, QC, by = "SampleId")
```

```{r echo=FALSE}
x = LUAD_all$Ploidy_ABS
y = LUAD_all$Ploidy_tumor_only

absdiff <- abs(x - y)
fract.grid <- which(absdiff < 0.5)

LUAD_all$absdiff = "discordant"
LUAD_all$absdiff[fract.grid] = "concordant"
```


```{r echo=FALSE, fig.width = 7, fig.height = 4.5, eval=FALSE}
# ## Purity of ploidy outliers
# **[Left]** Purity of PureCN and ABSOLUTE results from ploidy-discordant samples 
# (the cyan dots from the above plot) are plotted. 
# **[Right]** Purity of PureCN and ABSOLUTE results from the ploidy-concordant samples. 

df = df[which(df$absdiff == "discordant"),]
x = "Purity_ABS"
y = "Purity_tumor_only"
plot_for = "purity"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "TCGA-LUAD Purity \n of discordant Ploidy")

source("~/data2/PureCN_manuscript/Figures/manuscript_figure.R")
discordant_puri = final_plot
rm(final_plot)

df = df_ploi
df = df[which(df$absdiff == "concordant"),]
x = "Purity_ABS"
y = "Purity_tumor_only"
plot_for = "purity"
customLabs = labs(x = "SNP6 array (ABSOLUTE)", 
                  y = "Tumor WES (PureCN)", 
                  title = "TCGA-LUAD Purity \n of cordant Ploidy")

source("~/data2/PureCN_manuscript/Figures/manuscript_figure.R")
concordant_puri = final_plot
rm(final_plot)

discordant_puri | concordant_puri
```



# Purity and Coverage
## OVC 
### Concordant
```{r}
selectedCol = c("Purity_tumor_only", "Ploidy_tumor_only", "mean.coverage.ontarget", "mean.duplication.ontarget")
summary(OV_all[which(OV_all$absdiff == "concordant"),][selectedCol])
```

### Discordant
```{r}
selectedCol = c("Purity_tumor_only", "Ploidy_tumor_only", "mean.coverage.ontarget", "mean.duplication.ontarget")
summary(OV_all[which(OV_all$absdiff == "discordant"),][selectedCol])
```

## LUAD 
### Concordant
```{r}
selectedCol = c("Purity_tumor_only", "Ploidy_tumor_only", "mean.coverage.ontarget", "mean.duplication.ontarget")
summary(LUAD_all[which(LUAD_all$absdiff == "concordant"),][selectedCol])
```

### Discordant
```{r}
selectedCol = c("Purity_tumor_only", "Ploidy_tumor_only", "mean.coverage.ontarget", "mean.duplication.ontarget")
summary(LUAD_all[which(LUAD_all$absdiff == "discordant"),][selectedCol])
```


# P-value
Significance of purity and coverage differences between concordant vs. discordant 
samples, where 'discordant' is defined as the ploidy difference >= 0.5.

## Purity
### OVC
```{r}
con_ind = which(OV_all$absdiff == "concordant")
dis_ind = which(OV_all$absdiff == "discordant")

wilcox.test(OV_all$Purity_tumor_only[con_ind],
            OV_all$Purity_tumor_only[dis_ind])
t.test(OV_all$Purity_tumor_only[con_ind], OV_all$Purity_tumor_only[dis_ind])
```

### LUAD
```{r}
con_ind = which(LUAD_all$absdiff == "concordant")
dis_ind = which(LUAD_all$absdiff == "discordant")

wilcox.test(LUAD_all$Purity_tumor_only[con_ind],
            LUAD_all$Purity_tumor_only[dis_ind])
t.test(LUAD_all$Purity_tumor_only[con_ind], LUAD_all$Purity_tumor_only[dis_ind])
```

### Both
```{r}
all = rbind(OV_all, LUAD_all)
con_ind = which(all$absdiff == "concordant")
dis_ind = which(all$absdiff == "discordant")

wilcox.test(all$Purity_tumor_only[con_ind],
            all$Purity_tumor_only[dis_ind])
t.test(all$Purity_tumor_only[con_ind], all$Purity_tumor_only[dis_ind])
```

## Coverage
### OVC
```{r}
con_ind = which(OV_all$absdiff == "concordant")
dis_ind = which(OV_all$absdiff == "discordant")

wilcox.test(OV_all$mean.coverage.ontarget[con_ind],
            OV_all$mean.coverage.ontarget[dis_ind])
t.test(OV_all$mean.coverage.ontarget[con_ind], OV_all$mean.coverage.ontarget[dis_ind])
```

### LUAD
```{r}
con_ind = which(LUAD_all$absdiff == "concordant")
dis_ind = which(LUAD_all$absdiff == "discordant")

wilcox.test(LUAD_all$mean.coverage.ontarget[con_ind],
            LUAD_all$mean.coverage.ontarget[dis_ind])
t.test(LUAD_all$mean.coverage.ontarget[con_ind], LUAD_all$mean.coverage.ontarget[dis_ind])
```

### Both
```{r}
all = rbind(OV_all, LUAD_all)
con_ind = which(all$absdiff == "concordant")
dis_ind = which(all$absdiff == "discordant")

wilcox.test(all$mean.coverage.ontarget[con_ind],
            all$mean.coverage.ontarget[dis_ind])
t.test(all$mean.coverage.ontarget[con_ind], all$mean.coverage.ontarget[dis_ind])
```