---
title: "CNVworkflow Figures"
author: "Sehyun Oh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, eval=FALSE}
if (!"pathwork" %in% installed.packages())
    devtools::install_github("thomasp85/patchwork")

suppressPackageStartupMessages({
  library(patchwork)
  library(pdftools)
})
```

```{r}
library(patchwork)
library(pdftools)
```

## Figure 1
* File under `/data2/PureCN_manuscript/Figures`   
    + F1A/B: `manuscript_puri_ploi.Rmd`
    + F1G: `manuscript_loh_t_plots.Rmd`
* File under `/data2/PureCN_manuscript/luad/Figures`
    + F1C/D: `luad_manuscript_puri_ploi.Rmd`
    + F1H: `luad_manuscript_loh_t_plots.Rmd`
* File under `/data2/PureCN_manuscript/Revision/LOH/script`
    + F1E: `LOH_bubble_plot_OVC.Rmd`
    + F1F: `LOH_bubble_plot_LUAD.Rmd`
    
```{r}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"
F1A = readRDS(file.path(data_dir, "Fig1A.ovc_puri.rds"))
F1B = readRDS(file.path(data_dir, "Fig1B.ovc_ploi.rds"))
F1C = readRDS(file.path(data_dir, "Fig1C.luad_puri.rds"))
F1D = readRDS(file.path(data_dir, "Fig1D.luad_ploi.rds"))

F1E = readRDS(file.path(data_dir, "Fig1E.ovc_loh_all.rds"))
# F1F = readRDS(file.path(data_dir, "Fig1F.luad_loh_all.rds"))
F1F = readRDS(file.path(data_dir, "Fig1F.luad_loh_all_v2.rds"))

F1G = readRDS(file.path(data_dir, "Fig1G.ovc_loh.rds"))
F1H = readRDS(file.path(data_dir, "Fig1H.luad_loh.rds"))
```

```{r fig.width=8, fig.height=9}
(F1A | F1B ) / 
  (F1C | F1D)
```

```{r fig.width=9, fig.height=5}
F1E | F1F
```

```{r fig.width=9.5, fig.height=3.5}
F1G | F1H
```

## Figure 2
* File under `/data2/PureCN_manuscript/Figures`   
    + F2A/E/F: `manuscript_benchmark_plots.Rmd`
    + F2C/D: `manuscript_benchmark_both.Rmd`
* File under `/Users:/sehyunoh/wallabe4_backup/github/PureCN_manuscript/luad/Figures`
    + F2B/G/H: `luad_manuscript_benchmark_plots.Rmd`

```{r}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"

F2A = readRDS(file.path(data_dir, "Fig2A.ovc_AUC_gain.rds"))
F2B = readRDS(file.path(data_dir, "Fig2B.luad_AUC_gain.rds"))

F2C = readRDS(file.path(data_dir, "Fig2C.ov_callrate_plot.rds"))
F2D = readRDS(file.path(data_dir, "Fig2D.luad_callrate_plot.rds"))

F2E = readRDS(file.path(data_dir, "Fig2E.ov_accuracy.rds"))
F2F = readRDS(file.path(data_dir, "Fig2F.luad_accuracy.rds"))
F2G = readRDS(file.path(data_dir, "Fig2G.ov_sgz_accuracy.rds"))
F2H = readRDS(file.path(data_dir, "Fig2H.luad_sgz_accuracy.rds"))
```

```{r warning = FALSE, fig.width=16, fig.height=9}
(F2A | F2B | F2C | F2D) /
  (F2E | F2F | F2G | F2H)
```

## Figure 3
* File under `/data2/PureCN_manuscript/Figures`   
    + F3A: `manuscript_TMB_both.Rmd`
    + F3B: `manuscript_mutSig_plots.Rmd`
* File under `/Users:/sehyunoh/wallabe4_backup/github/PureCN_manuscript/luad/Figures`
    + F3C: `luad_manuscript_mutSig_plots.Rmd`
    
```{r}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"

F3A = readRDS(file.path(data_dir, "Fig3A.TMB.rds"))
F3B = readRDS(file.path(data_dir, "Fig3B.ovc_mutSig.rds"))
F3C = readRDS(file.path(data_dir, "Fig3C.luad_mutSig.rds"))
```

```{r fig.width=13.5, fig.height=5.4, eval=FALSE, echo=FALSE}
F3A + 
  (F3B + F3C + plot_layout(nrow = 1, width = c(1, 2))) + 
  plot_layout(nrow = 1, width = c(2, 3.5))
```

```{r fig.width=4.1, fig.height=4.5}
F3A
```

```{r fig.width=8, fig.height=3.5}
F3B + F3C + plot_layout(nrow = 1, width = c(1, 2))
```

## Sup. Figure 2
* File under `/data2/FACETS`   
    + SF2A/B: `OVC/Results/Figure_FACETS.Rmd`
    + SF2C/D: `LUAD/Results/Figure_FACETS_luad.Rmd`

```{r}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"

SF2A = readRDS(file.path(data_dir, "SFig2A.ovc_puri.rds"))
SF2B = readRDS(file.path(data_dir, "SFig2B.ovc_ploi.rds"))
SF2C = readRDS(file.path(data_dir, "SFig2C.luad_puri.rds"))
SF2D = readRDS(file.path(data_dir, "SFig2D.luad_ploi.rds"))
```

```{r fig.width=7.5, fig.height=9.3}
(SF2A | SF2B) / (SF2C | SF2D)
```

```{r fig.width=7.5, fig.height=4.7, eval=FALSE, echo=FALSE}
(SF2C | SF2D) 
```

## Sup. Figure 3
* File under `/data2/PureCN_manuscript/Figures`   
    + SF3: `manuscript_concordance_plot.Rmd`
```{r fig.width=8, fig.height= 5}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"

SF3 = readRDS(file.path(data_dir, "SFig3.concordance.rds"))
SF3
```

## Sup. Figure 4
* File under `/data2/PureCN_manuscript/Figures`   
    + SF4A/B: `manuscript_benchmark_both.Rmd`

```{r}
data_dir = "/data2/PureCN_manuscript/Figures/Final_Figures"

SF4A = readRDS(file.path(data_dir, "SFig4A.cov_plot_purecn.rds"))
SF4B = readRDS(file.path(data_dir, "SFig4B.cov_plot_sgz.rds"))
```

```{r fig.width = 6, fig.height = 5.5}
SF4A
SF4B
```

# PDF
```{r Fig1_pdf, eval=FALSE}
pdf(file.path(data_dir, "Figure_1A_D.pdf"), 8, 9)
(F1A | F1B ) / 
  (F1C | F1D)
dev.off()

pdf(file.path(data_dir, "Figure_1E_F.pdf"), 9, 5)
(F1E | F1F)
dev.off()

pdf(file.path(data_dir, "Figure_1G_H.pdf"), 9.5, 3.5)
(F1G | F1H)
dev.off()

pdf_combine(c("Figure_1A_D.pdf", "Figure_1E_F.pdf", "Figure_1G_H.pdf"), output = "Figure_1.pdf")
```

```{r Fig2_pdf}
pdf(file.path(data_dir, "Figure_2.pdf"), 16, 9)
(F2A | F2B | F2C | F2D) /
  (F2E | F2F | F2G | F2H)
dev.off()
```

```{r Fig3_pdf}
pdf(file.path(data_dir, "Figure_3A.pdf"), 4.1, 4.5)
F3A
dev.off()

pdf(file.path(data_dir, "Figure_3B_C.pdf"), 8, 3.5)
F3B + F3C + plot_layout(nrow = 1, width = c(1, 2))
dev.off()

pdf_combine(c("Figure_3A.pdf", "Figure_3B_C.pdf"), output = "Figure_3.pdf")
```

```{r SFig2_pdf}
pdf(file.path(data_dir, "Supplementary_Figure_2.pdf"), 7.5, 9.3)
(SF2A | SF2B) / (SF2C | SF2D)
dev.off()
```

```{r SFig3_pdf}
pdf(file.path(data_dir, "Supplementary_Figure_3.pdf"), 8, 5)
SF3
dev.off()
```

```{r SFig4_pdf}
pdf(file.path(data_dir, "Supplementary_Figure_4.pdf"), 6, 5.5)
SF4A
SF4B
dev.off()
```
